---
title: "DE analysis for transcripts"
author:
- tabbassidaloii
output:
 md_document:
    variant: markdown_github
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

'Note:' The muscle abbreviations were GRA(for Gracilis), SED(semitendinosus distal),  SEM(semitendinosus middle)), GAL(gastrocnemius lateralis), REF(rectus femoris), VAM(vastus medialis) and VAL(vastus lateralis) in all the analyses(scripts) and they were only changed to GR, STM, STD, GL, RF, VM and VL for the plots(shown in the publication).

```{r setup, include = FALSE}
rm(list = ls())
if(!require("BiocManager", quietly = TRUE))install.packages("BiocManager")
if(!"limma" %in% installed.packages())BiocManager::install("limma")
if(!"edgeR" %in% installed.packages())BiocManager::install("edgeR")
if(!"lmerTest" %in% installed.packages())install.packages("lmerTest")
if(!"lsmeans" %in% installed.packages())install.packages("lsmeans")
if(!"dplyr" %in% installed.packages())install.packages("dplyr")
if(!"tibble" %in% installed.packages())install.packages("tibble")
if(!"knitr" %in% installed.packages())install.packages("knitr")
if(!"ggplot2" %in% installed.packages())install.packages("ggplot2")
if(!"reshape2" %in% installed.packages())install.packages("reshape2")

suppressPackageStartupMessages({ 
  library(edgeR)
  library(limma)
  library(lmerTest)
  library(lsmeans)
  library(dplyr)
  library(tibble)
  library(knitr)
  library(ggplot2)
  })

opts_chunk$set(eval = TRUE, echo = TRUE, tidy = TRUE, highlight = TRUE, dev = c('png'), fig.width = 10, fig.height = 8,  fig.path = "figures/")

options(scipen = 999)

col.muscle <- c("GAL" = "#999999", "GRA" = "#0099FF", "REF" = "#FF9999", 
                "SED" = "#0033FF", "SEM" = "#0066FF", "VAL" = "#FF6666", 
                "VAM" = "#FF3333")

setwd(dirname(rstudioapi::getSourceEditorContext()$path))
# loading data
load("outputs/correctedTranscriptsDatasetToAnalysis.RData")
```

## Differential expression analysis
We used the linear mixed models considering individual as a random effect, and muscle and batch as fixed effects.

The lmer function form lmerTest R package is used. And to find if there is any diffrence between muscles or batches the anova is performed. For the pairwise comparison we used Least-Squares Means(lsmeants R function)

```{r DEA, echo = FALSE, warning = FALSE}
if(file.exists("outputs/DE/DEA_transcripts.RData")){
  load("outputs/DE/DEA_transcripts.RData")} else{
    # Creating a design matrix 
    design <- model.matrix(~ group + individual, data = D_BatchMus$samples)
    colnames(design) <- gsub("group|individual", "", colnames(design))
    # mean-variance trend was removed using the voom
    v_BatchMus <- voom(D_BatchMus, design, plot = TRUE)
    DEtest_BatchMus <- matrix(data = NA, nrow = nrow(v_BatchMus$E), ncol = 140)
    rownames(DEtest_BatchMus) <- rownames(v_BatchMus$E)
    AllPossPairs <- combn(1 : length(names(col.muscle)), 2, simplify = FALSE)
    colnames(DEtest_BatchMus)[112:140] <- c("AveExpr", paste0("AveExpr", sort(names(col.muscle))),
                                            paste0("Log2FC_", unlist(lapply(AllPossPairs, function(x)
                                              paste(names(col.muscle[x]), collapse = "_"))))) 
    DEfixef_BatchMus <-  matrix(data = NA, nrow = nrow(v_BatchMus$E), ncol = 7)
    rownames(DEfixef_BatchMus) <- rownames(v_BatchMus$E)
    # fit linear-mixed model for each transcript (including individual as random and all factors as fixed effects)
    # RIN is excluded as we have missing value for 12 samples
    
    for(transcript in 1:nrow(v_BatchMus$E)){
      # Data
      DE.Data <- cbind(v_BatchMus$E[transcript,], v_BatchMus$weights[transcript,], v_BatchMus$targets)
      colnames(DE.Data)[1:2]<- c("values", "weights")
      # fit LMM 
      MM <- lmerTest::lmer(values ~ group +(1|individual), weights = weights, data = DE.Data)
      # Muscle
      # Pairwise comparison 
      PairWiseComp <- as.data.frame(lsmeans::lsmeans(MM, pairwise ~ group)$contrast)
      PairWiseComp <- reshape2::melt(PairWiseComp, id = "contrast"); 
      PairWiseComp <- PairWiseComp[order(PairWiseComp$contrast),]
      DEtest_BatchMus[transcript, 1:111] <- c(unlist(anova(MM)["group",]), PairWiseComp$value)
      # Extract the fixed-effect coefficients using fixef()
      DEfixef_BatchMus[transcript,] <- lme4::fixef(MM)
      # Average expression
      DEtest_BatchMus[transcript, 112] <- mean(DE.Data$values)
      # Average for muscle
      meanMus <- data.frame(DE.Data %>% group_by(group) %>% summarise(mean(values)))
      meanMus <- meanMus[sort(meanMus$group),]
      DEtest_BatchMus[transcript, c(113:119)] <- meanMus[sort(meanMus$group), 2]
      # Log2FC
      logFC <- c()
      for(i in 1:length(AllPossPairs)){
        p <- AllPossPairs[[i]]
        logFC[i] <- meanMus[p[1],2] - meanMus[p[2],2]
      }
      DEtest_BatchMus[transcript, c(120:140)] <- logFC
    }
    DEtest_BatchMus <- as.data.frame(DEtest_BatchMus)
    colnames(DEtest_BatchMus)[1:111] <- c(colnames(anova(MM)["group",]),
                                          paste(PairWiseComp$contrast, PairWiseComp$variable))
    DEfixef_BatchMus <- as.data.frame(DEfixef_BatchMus)
    colnames(DEfixef_BatchMus) <- names(lme4::fixef(MM))
    # I see to warning messages:
    ##(1) boundary(singular) fit: see ?isSingular
    # We have this warning when the Std.Dev. for random effect is equal to zero
    #(https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#singular-models-random-effect-variances-estimated-as-zero-or-correlations-estimated-as---1)
    ##(2) In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
    ##Model failed to converge with max|grad| = 0.0033159(tol = 0.002, component 1)
    # We have this warning when the random effect(individual) is explaining everything and leaving novariation at all for either any fixed effect or the residual variance.
    #(https://stackoverflow.com/questions/53034261/warning-lme4-model-failed-to-converge-with-maxgrad)

    DEtest_BatchMus$`FDR(>F)` <- p.adjust(DEtest_BatchMus[,"Pr(>F)"], method = "fdr")

    # Saving the dataset for the analysis
    save(v_BatchMus, DEtest_BatchMus, DEfixef_BatchMus, file = "outputs/DE/DEA_transcripts.RData")
  }
rm(list = setdiff(ls(), c("v_BatchMus", "D_BatchMus", "DEtest_BatchMus", "DEfixef_BatchMus", "col.muscle")))
```

### The muscle effect
The anova is used to find the transcripts that show different expression levels considering the muscle type as covariate.
```{r DEmuscle, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 5, fig.width = 5}
# Subsetting the differentially expressed transcripts
DEtestSig <- DEtest_BatchMus %>% tibble::rownames_to_column() %>%
  filter(`FDR(>F)` < 0.05) %>%
  arrange(`FDR(>F)`) %>%
  tibble::column_to_rownames("rowname") 
print("Number of DE transcripts") 
nrow(DEtestSig)
```

We use lsmeans to have pairwise comparison between muscles(muscle type is the the covariate).
There are 21 different pairwise comparisons. In the below table, the first column is the number of pairwise comparison that the transcript is deferentially expressed and the second column is the number of transcripts  

```{r DEmusclePairwise, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 5, fig.width = 5}
print(knitr::kable(data.frame(table(rowSums(DEtestSig[, grep("p.val", colnames(DEtestSig))]< 0.05)))))
```

The below plot shows the number of deferentially expressed transcripts(FDR < 0.05) in the pairwise comparisons of all the muscles.
The color intensity indicates the number of DE transcripts(the darker the color, the higher the number).

```{r DEmuscleTable, echo = FALSE, warning = FALSE, message = FALSE}
# Counting the proportion of differentially expressed transcripts in each pairwise comparisons
ColPVAl <- colnames(DEtest_BatchMus)[grep(" p.value", colnames(DEtest_BatchMus))]
MusDissSim <- matrix(0, nrow = 7, ncol = 7) 
colnames(MusDissSim) <- rownames(MusDissSim) <- names(col.muscle) 
for(c in ColPVAl){
  MusDissSim[gsub(".*- | p.*", "", c), gsub(" -.*", "", c)] <- MusDissSim[gsub(" -.*", "", c), gsub(".*- | p.*", "", c)] <- table(DEtestSig[c]< 0.05)["TRUE"]*100/nrow(DEtest_BatchMus)
}
MusDissSim[is.na(MusDissSim)] = 0
```


```{r DEmusclePlots1, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 5, fig.width = 5}
# New Abbreviations
muscleAbbreviation <- c("GAL" = "GL", "GRA" = "GR", "VAL" = "VL", "VAM" = "VM", "SED" = "STD", "SEM" = "STM", "REF" = "RF")
colnames(MusDissSim) <- muscleAbbreviation[match(colnames(MusDissSim), names(muscleAbbreviation))]
rownames(MusDissSim) <- muscleAbbreviation[match(rownames(MusDissSim), names(muscleAbbreviation))]

(Tosave <- pheatmap::pheatmap(MusDissSim, 
                              color = colorRampPalette(c("white", "gray"))(10),
                              scale = "none", number_format = "%.1f", 
                              cluster_cols = T, cluster_rows = T, 
                              fontsize = 12, key =F, angle_col = 90, 
                              legend_breaks = c(0, 1, 2, 3, 4),
                              display_numbers = T, na_col = "white"))
ggsave(Tosave, device = "tiff", units = "cm", width = 10, height = 7, filename = "figures/DE_transcript.tif")

# Removing transcripts originating from cell type related genes
load("outputs/cellTypeComposition/geneCellTypesEigengeneCor.RData")
geneCellTypeRelated <- unique(unlist(SigGenesCellTypes))
transcriptCellTypeRelated <- v_BatchMus$genes$ensembl_transcript_id[v_BatchMus$genes$ensembl_gene_id %in% geneCellTypeRelated] 
DEtestSigNotCellRelated <- DEtestSig[!rownames(DEtestSig) %in% transcriptCellTypeRelated,]

# Counting the proportion of differentially expressed transcripts (not cell type related) in each pairwise comparisons
ColPVAl <- colnames(DEtest_BatchMus)[grep(" p.value", colnames(DEtest_BatchMus))]
MusDissSim <- matrix(0, nrow = 7, ncol = 7) 
colnames(MusDissSim) <- rownames(MusDissSim) <- names(col.muscle) 
for(c in ColPVAl){
  MusDissSim[gsub(".*- | p.*", "", c), gsub(" -.*", "", c)] <- MusDissSim[gsub(" -.*", "", c), gsub(".*- | p.*", "", c)] <- table(DEtestSigNotCellRelated[c]< 0.05)["TRUE"]*100/nrow(DEtest_BatchMus)
}
MusDissSim[is.na(MusDissSim)] = 0
colnames(MusDissSim) <- muscleAbbreviation[match(colnames(MusDissSim), names(muscleAbbreviation))]
rownames(MusDissSim) <- muscleAbbreviation[match(rownames(MusDissSim), names(muscleAbbreviation))]

(Tosave <- pheatmap::pheatmap(MusDissSim, 
                              color = colorRampPalette(c("white", "gray"))(10),
                              scale = "none", number_format = "%.1f", 
                              cluster_cols = T, cluster_rows = T, 
                              fontsize = 12, key =F, angle_col = 90, 
                              legend_breaks = c(0, 0.5, 1, 1.5),
                              display_numbers = T, na_col = "white", ))
ggsave(Tosave, device = "tiff", units = "cm", width = 10, height = 7, filename = "figures/DE_transcript_not_cell_type_related.tif")
```

The GAL is the most different muscle.
The VAM, REF and VAL are similar and clustered together and GRA, SEM and SED are also more similar as they also cluster together.

```{r DEsummary1, echo = FALSE, warning = FALSE, message = FALSE}
# Summary of DE results
contrasts <- gsub(" p.value", "", colnames(DEtest_BatchMus)[grep("p.value",colnames(DEtest_BatchMus))])
for(i in contrasts){
  M1 <- gsub(" -.*", "", i)
  M2 <- gsub(".*- ", "", i)
  LogCol <- colnames(DEtest_BatchMus)[grep(paste0("Log2FC", "_", M1, "_", M2),colnames(DEtest_BatchMus))]
  PvalCol <- colnames(DEtest_BatchMus)[grep(paste0(i, " p.value"),colnames(DEtest_BatchMus))]
  HigherInM1 <- rownames(DEtestSig[DEtestSig[PvalCol] < 0.05 & DEtestSig[LogCol] > 0,])
  HigherInM2 <- rownames(DEtestSig[DEtestSig[PvalCol] < 0.05 & DEtestSig[LogCol] < 0,])
  assign(paste0("HigherIn", M1, "than", M2), HigherInM1)
  assign(paste0("HigherIn", M2, "than", M1), HigherInM2)
  rm(HigherInM1, HigherInM2)
}

summaryDETbale <- data.frame(Comparison = ls(pattern = paste0("HigherIn.*", "$")),
                             NoGenes = unlist(lapply(ls(pattern = paste0("HigherIn.*", "$")),
                                                     function(x) length(get(x)))))
print(knitr::kable(head(summaryDETbale)))
```

Below plot visualizes the overlap of higher expressed transcripts than any other muscles.
There are many transcripts that are only higher in GAL than one or more other muscles
```{r musclePlots2, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 6, fig.width = 6}
# Transcripts higher expressed in each muscle in comparison with any other muscle
for(Mus in names(col.muscle)){
    TranscriptsHighInMus <- unique(unlist(lapply(ls(pattern = paste0("HigherIn", Mus, ".*", "$")), get)))
    assign(paste0("TranscriptsHighIn", Mus), TranscriptsHighInMus)
    rm(TranscriptsHighInMus)
}

# Transcripts that are higher at least in one muscle
TranscriptsHigh <- unique(unlist(lapply(ls(pattern = paste0("TranscriptsHighIn.*", "$")), get)))
AllSets <- matrix(0, ncol = 7, nrow = length(TranscriptsHigh))
rownames(AllSets) <- TranscriptsHigh
colnames(AllSets) <- ls(pattern = paste0("TranscriptsHighIn.*", "$"))
for(set in ls(pattern =  paste0("TranscriptsHighIn.*", "$"))){
  AllSets[,set] <- ifelse(rownames(AllSets) %in% get(set), 1, 0)
}

pheatmap::pheatmap(AllSets, color = c("grey", "darkblue"),
                   legend = F, show_rownames = F, treeheight_row = 0)
```
  
### Heatmap of log-CPM values for 1000 top differentially expressed transcripts
```{r musclePlots4, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 14, fig.width = 7}
# Heatmap for top DE transcripts 
# Get the transcripts for the top 1000 most DE transcripts
HighDE <- rownames(DEtestSig[order(DEtestSig$`FDR(>F)`),])[1:1000]
# Subset matrix
lcpm <- cpm(D_BatchMus, log = T)
hmDat <- lcpm[HighDE,]
# Set the colors
col.batch <- c("1" = "#e6194B", "2" = "#3cb44b", "3" = "#4363d8", "4" = "#911eb4", "5" = "black")
BatchColPlot <- col.batch[match(D_BatchMus$samples$batch[match(colnames(D_BatchMus), 
                                                               rownames(D_BatchMus$samples))], 
                                names(col.batch))]
MusColPlot <- col.muscle[match(gsub("MD[0-9][0-9]_|_Sam.*", "", colnames(D_BatchMus)), 
                                  names(col.muscle))]
col.individual <- c("MD06" = "#A6CEE3", "MD07" = "#1F78B4", "MD08" = "#B2DF8A", 
                     "MD10" = "#33A02C", "MD11" = "#FB9A99", "MD12" = "#E31A1C", 
                     "MD13" = "#FDBF6F", "MD14" = "#FF7F00", "MD15" = "#66C2A5", 
                     "MD16" = "#FC8D62", "MD17" = "#8DA0CB", "MD18" = "#E78AC3", 
                     "MD19" = "#A6D854", "MD20" = "#FFD92F", "MD21" = "#E5C494", 
                     "MD22" = "#B3B3B3", "MD26" = "#1B9E77", "MD27" = "#D95F02", 
                     "MD28" = "#7570B3", "MD31" = "#E7298A")
IndColPlot <- col.individual[match(gsub("_.*", "", colnames(D_BatchMus)), 
                                  names(col.individual))]
myCols <- cbind(BatchColPlot, MusColPlot, IndColPlot)
colnames(myCols) <- c("Batch", "Muscle", "Individual")
# Plot the heatmap
heatmap3::heatmap3(hmDat, ColSideColors = myCols,
              labCol = gsub("_Sam.*", "", colnames(hmDat)), 
              main = "Top 1000 DE transcripts", 
              labRow = FALSE, scale = "row",
              col =   gplots::colorpanel(1000, "blue", "white", "red"),
              margins = c(5,1), Rowv = T, showRowDendro = F)
```

```{r saveSelectedDataset, echo = FALSE, message = FALSE, warning = FALSE}
# colnames(DEtest_BatchMus) = gsub("GAL", "GL", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("GRA", "GR", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("SEM", "STM", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("SED", "STD", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("VAM", "VM", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("VAL", "VL", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("REF", "RF", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub(" - ", "-", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_GR", "GL-GR Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_RF", "GL-RF Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_STD", "GL-STD Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_STM", "GL-STM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_VL", "GL-VL Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GL_VM", "GL-VM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GR_RF", "GR-RF Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GR_STD", "GR-STD Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GR_STM", "GR-STM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GR_VL", "GR-VL Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_GR_VM", "GR-VM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_RF_STD", "RF-STD Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_RF_STM", "RF-STM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_RF_VL", "RF-VL Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_RF_VM", "RF-VM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_STD_STM", "STD-STM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_STD_VL", "STD-VL Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_STD_VM", "STD-VM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_STM_VL", "STM-VL Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_STM_VM", "STM-VM Log2FC", colnames(DEtest_BatchMus))
# colnames(DEtest_BatchMus) = gsub("Log2FC_VL_VM", "VL-VM Log2FC", colnames(DEtest_BatchMus))
# 
# DEtest_BatchMus <- DEtest_BatchMus %>%
#   tibble::rownames_to_column() %>%
#   rename(`ensembl transcript id` = rowname) %>%
#   select (`ensembl transcript id`, `F value`, `Pr(>F)`, `FDR(>F)`, `GL-GR Log2FC`, `GL-GR t.ratio`, `GL-GR p.value`, `GL-RF Log2FC`, `GL-RF t.ratio`, `GL-RF p.value`, `GL-STD Log2FC`, 
# `GL-STD t.ratio`, `GL-STD p.value`, `GL-STM Log2FC`, `GL-STM t.ratio`, `GL-STM p.value`, `GL-VL Log2FC`, `GL-VL t.ratio`,  
# `GL-VL p.value`, `GL-VM Log2FC`, `GL-VM t.ratio`, `GL-VM p.value`, `GR-RF Log2FC`, `GR-RF t.ratio`, `GR-RF p.value`, 
# `GR-STD Log2FC`, `GR-STD t.ratio`, `GR-STD p.value`, `GR-STM Log2FC`, `GR-STM t.ratio`, `GR-STM p.value`, `GR-VL Log2FC`,
# `GR-VL t.ratio`, `GR-VL p.value`, `GR-VM Log2FC`, `GR-VM t.ratio`, `GR-VM p.value`, `RF-STD Log2FC`, `RF-STD t.ratio`, 
# `RF-STD p.value`, `RF-STM Log2FC`, `RF-STM t.ratio`, `RF-STM p.value`, `RF-VL Log2FC`, `RF-VL t.ratio`, `RF-VL p.value`, 
# `RF-VM Log2FC`, `RF-VM t.ratio`, `RF-VM p.value`, `STD-STM Log2FC`, `STD-STM t.ratio`, `STD-STM p.value`, `STD-VL Log2FC`,  
# `STD-VL t.ratio`, `STD-VL p.value`, `STD-VM Log2FC`, `STD-VM t.ratio`, `STD-VM p.value`, `STM-VL Log2FC`, `STM-VL t.ratio`, 
# `STM-VL p.value`, `STM-VM Log2FC`, `STM-VM t.ratio`, `STM-VM p.value`, `VL-VM Log2FC`, `VL-VM t.ratio`, `VL-VM p.value`, 
# )
# DEtest_BatchMus[, -c(1)] <- apply(DEtest_BatchMus[, -c(1)], 2, function (x) round(x, digits = 5)) 
# 
# DEtest_BatchMus %>% write.csv("outputs/DE/DEt_ranscripts.csv",
#                           quote = F, row.names = F)
```

```{r}
sessionInfo()
```
