---
title: "WGCNA"
author:
- tabbassidaloii
output:
 md_document:
    variant: markdown_github
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

'Note:' The muscle abbreviations were GRA(for Gracilis), SED(semitendinosus distal),  SEM(semitendinosus middle)), GAL(gastrocnemius lateralis), REF(rectus femoris), VAM(vastus medialis) and VAL(vastus lateralis) in all the analyses(scripts) and they were only changed to GR, STM, STD, GL, RF, VM and VL for the plots(shown in the publication).

```{r setup, include = FALSE}
rm(list = ls())
if(!"limma" %in% installed.packages())BiocManager::install("limma") 
if(!"edgeR" %in% installed.packages())BiocManager::install("edgeR") 
if(!"WGCNA" %in% installed.packages())BiocManager::install("WGCNA") 
if(!"flashClust" %in% installed.packages())install.packages("flashClust")
if(!"dplyr" %in% installed.packages())install.packages("dplyr")
if(!"readxl" %in% installed.packages())install.packages("readxl")

suppressPackageStartupMessages({ 
  library(WGCNA)
  library(flashClust)  
  library(dplyr)
  library(readxl)
  })
opts_chunk$set(eval = TRUE, echo = TRUE, tidy = TRUE, highlight = TRUE, dev = c('png'), fig.width = 10, fig.height = 8,  fig.path = "figures/")

BiocStyle::markdown()
options(scipen = 999)

# Loading the data
Data <- new.env(); load("outputs/correctedGenesDatasetToAnalysis.RData", envir = Data)
D_BatchMus <- Data$D_BatchMus
# Creating a design matrix
design <- model.matrix(~ group + individual, data = D_BatchMus$samples)
colnames(design) <- gsub("group|individual", "", colnames(design))
# mean-variance trend was removed using the voom
v_BatchMus <- voom(D_BatchMus, design, plot = TRUE)

# Set colors
col.muscle <- c("GAL" = "#999999", "GRA" = "#0099FF", "REF" = "#FF9999", 
                "SED" = "#0033FF", "SEM" = "#0066FF", "VAL" = "#FF6666",  
                "VAM" = "#FF3333")
MusColPlot <- col.muscle [match(gsub("MD[0-9][0-9]_|_Sam.*", "", colnames(D_BatchMus)), names(col.muscle))]
col.batch <- c("1" = "#e6194B", "2" = "#3cb44b", "3" = "#4363d8", "4" = "#911eb4", "5" = "black")
BatchColPlot <- col.batch [match(D_BatchMus$samples$batch [match(colnames(D_BatchMus), rownames(D_BatchMus$samples))], names(col.batch))]
col.individual <- c("MD06" = "#A6CEE3", "MD07" = "#1F78B4", "MD08" = "#B2DF8A", 
                     "MD10" = "#33A02C", "MD11" = "#FB9A99", "MD12" = "#E31A1C", 
                     "MD13" = "#FDBF6F", "MD14" = "#FF7F00", "MD15" = "#66C2A5", 
                     "MD16" = "#FC8D62", "MD17" = "#8DA0CB", "MD18" = "#E78AC3", 
                     "MD19" = "#A6D854", "MD20" = "#FFD92F", "MD21" = "#E5C494", 
                     "MD22" = "#B3B3B3", "MD26" = "#1B9E77", "MD27" = "#D95F02", 
                     "MD28" = "#7570B3", "MD31" = "#E7298A")
IndColPlot <- col.individual [match(gsub("_.*", "", colnames(D_BatchMus)), names(col.individual))]
col.lane <- c("L003.1" = "lightblue", "L004.1" = "darkblue", "L002.2" = "pink", "L004.2" = "red")
LaneColPlot <- col.lane [match(D_BatchMus$samples$lane [match(colnames(D_BatchMus), rownames(D_BatchMus$samples))], names(col.lane))]
col.RNA <- c("WithKit" = "#bfef45", "WithOutKit" = "#f58231")
RNAColPlot <- col.RNA [match(D_BatchMus$samples$IsolationRNA [match(colnames(D_BatchMus), rownames(D_BatchMus$samples))], names(col.RNA))]
col.Hos <- c("EMC" = "green", "HMC" = "orange")
HosColPlot <- col.Hos [match(D_BatchMus$samples$Hospital [match(colnames(D_BatchMus), rownames(D_BatchMus$samples))], names(col.Hos))]
```

## Gene coexpression network analysis
### Finding the outlier samples using Eucidian distance
```{r OutlierSample, message = FALSE, warning = FALSE, fig.width = 14}
# show that row names agree	
all(rownames(v_BatchMus$targets) == colnames(v_BatchMus$E)) 	
# sample network based on squared Euclidean distance	
# note that data should be transposed 	
A <- WGCNA::adjacency(v_BatchMus$E, type = "distance", corFnc = WGCNA::cor)	
# this calculates the whole network connectivity	
k <- as.numeric(apply(A, 2, sum)) - 1	
# standardized connectivity	
Z.k <- scale(k)	
# Designate samples as outlying	
# if their Z.k value is below the threshold	
thresholdZ.k <- -2.5 # often -2.5	
# the color vector indicates outlyingness(red)	
outlierColor <- ifelse(Z.k < thresholdZ.k, "red", "black")	
# calculate the cluster tree using flahsClust or hclust	
sampleTree <- flashClust(as.dist(1 - A), method = "average")	
# Convert traits to a color representation where red indicates high values	
traitColors <- data.frame(Muscle = MusColPlot,
                           Individual = IndColPlot,
                           Hospital = HosColPlot,
                           RNAprotocol = RNAColPlot, 
                           Batch = BatchColPlot,
                           Lane = LaneColPlot)

traitColors <- cbind(traitColors, data.frame(numbers2colors(sapply(v_BatchMus$targets %>% dplyr::select("lib.size", "RIN", "Concen.", "age"), as.numeric), signed = FALSE)))
dimnames(traitColors) [[2]][7:10] <- c("LibSize", "RIN", "Concen.", "Age")
datColors <- data.frame(outlierC = outlierColor, traitColors)	
# Plot the sample dendrogram and the colors underneath. 	
plotDendroAndColors(sampleTree, dendroLabels = gsub("_Sam.*", "",colnames(v_BatchMus$E)), groupLabels = names(datColors), colors = datColors, main = "Sample dendrogram and trait heatmap")
# Remove outlying samples from expression and phenotypic data	
```

In general the sample dendrogram (based on Euclidean distance) confirms our earlier observation, bigger inter-individual difference than the muscle differences!
REF MD13 and VAM MD28 are detected as outlier samples but we don't need to exclude these samples since the dataset is big and these two samples will not have big effect on the network construction since the `corOptions = c("maxPOutliers = 0.1")` argument is also specified for the network construction.

In addition, these two samples are not that far away form other samples looking at the PCA plots.

### WGCNA in 5 steps (as explianed in Abbassi-Daloii, 2020, DOI: 10.1016/j.ygeno.2020.05.026)
1. Calculating the dissimilarity matrix using 7 different powers(6, 8, 10, 12, 14, 18, 22)
2. Detecting module using 3 settings for minClusterSize(15, 20, 30) and 3 settings for different deepSplit(0, 2, 4) 
3. Merging modules using 5 settings for cutHeight(0.1, 0.15, 0.2, 0.25, 0.3)
4. Defining co-expressed pairs(CPs)
5. Defining knowledge pairs(KPs)
6. Finding the overlap between KPs and CPs
7. Creating settings summary file

```{r WGCNA_power, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
enableWGCNAThreads()
## Choose a set of soft thresholding powers	
powers <- c(c(1 : 10), seq(12, 30, 2))	
# choose power based on SFT criterion	
# conflict_prefer("cor", "WGCNA")
# conflict_prefer("which.max", "raster")
# conflict_prefer("filter", "dplyr")
# conflict_prefer("paste", "base")
sft <- pickSoftThreshold(t(v_BatchMus$E), powerVector = powers, corFnc = bicor, 	
                           corOptions = c("maxPOutliers = 0.1"), networkType = "signed hybrid")	
# Plot the results: 	
par(mar = c(5, 4, 4, 2) + 0.1, mfrow = c(1, 2))	
# SFT index as a function of different powers	
plot(sft$fitIndices [, 1], - sign(sft$fitIndices [, 3]) * sft$fitIndices [, 2] ,	
       xlab = "Soft Threshold(power)", ylab = "Scale Free Topology Model Fit, signed R^2", type = "n", 
       main = "Scale independence")	
text(sft$fitIndices [, 1], - sign(sft$fitIndices [, 3]) * sft$fitIndices [, 2] ,	
labels = powers, col = "red")	
# R^2 cut-off	
abline(h = 0.8, col = "blue")	
# Mean connectivity as a function of different powers	
plot(sft$fitIndices [, 1], sft$fitIndices [, 5], type = "n", 	
       xlab = "Soft Threshold(power)", ylab = "Mean Connectivity", main = "Mean connectivity")
text(sft$fitIndices [, 1], sft$fitIndices [, 5], labels = powers, col = "red")	
```

The Powers that we selected for the network construction all have R2 above 0.8

```{r AllStepsWGCNA1, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
# 1. Adjacency and dissTOM
mainDir = "outputs/WGCNA/"
subDir = "dissTOM"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(length(list.files(paste0(mainDir, subDir))) == 0) {
  # Create correlation matrix
  A <- adjacency(t(v_BatchMus$E), power = 1, type = "signed hybrid", corFnc = "bicor", corOptions = c("maxPOutliers = 0.1"))
  # change the default power value in adjacency function(WGCNA) to 1
  adjacencyPower <- function(cor_mat, power) {abs(cor_mat) ^ power}
  # Weighted adjacency matrix, using different powers:
  power <- c(6, 8, 10, 12, 14, 18, 22)
  # Parallel computing considering different powers

  for(i in 1:length(power)) {
      # Raised adjacency matrix to different powers 
      Adjacency <- adjacencyPower(A, power = power[i])
      save(Adjacency, file = paste0("outputs/WGCNA/Adjacency/AdjacencyPower", power[i], ".RData"))
      # Define a dissimilarity based on the topological overlap
      TOM <- WGCNA::TOMsimilarity(Adjacency, TOMType = "signed")
      colnames(TOM) <- rownames(TOM) <- colnames(Adjacency)
      dissTOM <- 1 - TOM
      # Hierarchical gene clustering(according to the dissimilarity matrix)
      geneTree <- flashClust::flashClust(as.dist(dissTOM), method = "average")
      save(dissTOM, geneTree, file = paste0("outputs/WGCNA/dissTOM/dissTOMPower", power[i], ".RData"))
      rm(Adjacency, TOM, dissTOM, geneTree)
  }
}

# 2. Module Detection
mainDir = "outputs/WGCNA/"
subDir = "ModuleDetection"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(length(list.files(paste0(mainDir, subDir))) == 0) {
  filenames <- list.files("outputs/WGCNA/dissTOM/", full.names = T)
  for(file in filenames){
    args <-  c(9, file, 3, 15, 20, 30, 3, 0, 2, 4)
    # loading dissimilarity matrix
    power <- gsub("[^[:digit:],]", "", args[2])
    load(args[2]) 
    # Module detection
    # Set the minimum module size 
    args <- as.numeric(args[-2])
    minModuleSize <- args[3 :(args[2] + 2)]
    deepSplit <- args[7:(args[6] + 6)]
    Combin <- expand.grid(minModuleSize, deepSplit)
    print(Combin)
    for(i in 1: nrow(Combin)) {
      # Module detection by cutting branches
      dynamicMods <- dynamicTreeCut::cutreeDynamic(dendro = geneTree, distM = dissTOM, method = "hybrid", 
                                                      deepSplit = Combin[i, 2], pamRespectsDendro = FALSE,
                                                      minClusterSize = Combin[i, 1])
      # Convert labels to colors for plotting
      dynamicColors <- WGCNA::labels2colors(dynamicMods)
      # Calculate eigengenes(clustring modules based on expression similarities)
      MEList <- WGCNA::moduleEigengenes(t(v_BatchMus$E), colors = dynamicColors)
      MEs <- MEList$eigengenes
      NAMES <- colnames(dissTOM)
      save(dynamicColors, geneTree, NAMES, file = paste0("outputs/WGCNA/ModuleDetection/ModuleDetectionPower", power,
                                                           "MinModuleSize", Combin[i, 1],
                                                           "deepSplit", Combin[i, 2], ".RData"))
    }
  }
}

# 3. Merge Modules
mainDir = "outputs/WGCNA/"
subDir = "MergeModules"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(length(list.files(recursive = TRUE, pattern = "All_genes_modules_Power", path = paste0(mainDir, subDir))) == 0) {
  filenames <- list.files("outputs/WGCNA/ModuleDetection/", full.names = T)
  SAVEPATH <- "outputs/WGCNA/MergeModules/"
  for(file in filenames){
    args <-  c(file, 5, 0.1, 0.15, 0.20, 0.25, 0.30)
    Setting <- gsub(".*ModuleDetection|.RData", "", args[1])
    print(Setting)
    # loading Modules
    load(args[1])
    # Merging similar modules
    args <- as.numeric(args[-1])
    CutHeight <- args[2 :(args[1] + 1)]
    for(i in 1: length(CutHeight)) { 
      # Call an automatic merging function
      merge <- WGCNA::mergeCloseModules(t(v_BatchMus$E), dynamicColors, cutHeight = CutHeight[i],
                                           verbose = 3, corFnc = bicor) 
      # The merged module colors
      mergedColors <- merge$colors
      # Eigengenes of the new merged modules:
      mergedMEs <- merge$newMEs
      write.csv(mergedMEs, paste0(SAVEPATH, "ModuleEigengenes_merged_", 
                                    Setting, "CutHeight", CutHeight[i], ".csv"), row.names = FALSE)
      # Summary output of network analysis results(after merging)
      module_colors <- unique(mergedColors)
      All_genes_modules <- as.data.frame(cbind(NAMES, mergedColors)) 
      # colnames(All_genes_modules) <- NULL
      write.csv(All_genes_modules, paste0(SAVEPATH, "All_genes_modules_", Setting,
                                            "CutHeight", CutHeight[i], ".csv"), row.names = FALSE)
      # names(colors) of the modules
      nSamples <- nrow(t(v_BatchMus$E))
      modNames <- substring(names(mergedMEs), 3)
      geneModuleMembership <- as.data.frame(bicor(t(v_BatchMus$E), mergedMEs, maxPOutliers = 0.1)) 
      MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples)) 
      names(geneModuleMembership) <- gsub("ME", "MM", names(geneModuleMembership))
      names(MMPvalue) <-  gsub("ME", "p.MM", names(MMPvalue))
      write.csv(merge(geneModuleMembership, MMPvalue, by=0, all = TRUE, sort = FALSE),
                 paste0(SAVEPATH, "GeneModule_Membership_MMPvalue_", Setting,
                         "CutHeight", CutHeight[i], ".csv"), col.names = TRUE, row.names = FALSE)
      phenWGCNA <- v_BatchMus$targets %>%
        mutate(group = as.numeric(group)) %>%
        mutate(individual = as.numeric(gsub("MD", "", individual))) %>%
        select(-"sampleID") %>%
        mutate(IsolationRNA = ifelse(IsolationRNA == "WithOutKit", 1, 0)) %>%
        mutate(Hospital = ifelse(Hospital == "HMC", 1, 0)) %>%
        mutate(lane = as.numeric(lane)) %>%
        mutate(batch = as.numeric(batch))
    
    
      # Module membership values and gene significance
      for(j in 1 : length(phenWGCNA)[1]) 
        {
        # Define interested variable of datTrait
        traitOfInterest <- as.data.frame(phenWGCNA [, j])
        names(traitOfInterest) <-  colnames(phenWGCNA)[j]
        ## Relating modules to external clinical traits and identifying important genes
        nGenes <- ncol(t(v_BatchMus$E))
        nSamples <- nrow(t(v_BatchMus$E))
        moduleTraitCor <- matrix(nrow = dim(mergedMEs)[2], ncol = dim(phenWGCNA)[2]) 
        colnames(moduleTraitCor) <- paste0("Cor_", colnames(phenWGCNA))
        rownames(moduleTraitCor) <- colnames(mergedMEs) 
        for(trait in colnames(phenWGCNA)) {
          moduleTraitCor [, grep(trait, colnames(moduleTraitCor))] <-
            cor(mergedMEs, phenWGCNA [, trait], use = "pairwise.complete.obs")
        }
        moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples)
        colnames(moduleTraitPvalue) <- paste0("Pvalue_", colnames(moduleTraitPvalue))
    
        # Writing the tables down with modules and correlation
        write.csv(merge(moduleTraitCor, moduleTraitPvalue, by = 0, all = TRUE, sort = FALSE),
                   paste0(SAVEPATH, "ModuleTrait_Cor_Pvalue_", Setting, "CutHeight",  CutHeight[i], ".csv"), 
                   col.names = TRUE, row.names = FALSE) 
    
    
        # correlation matrix of each gene and each traits
        geneTraitSignificance <- as.data.frame(cor(t(v_BatchMus$E), traitOfInterest,
                                                use = 'pairwise.complete.obs'))
        GSPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples)) 
        names(geneTraitSignificance) <- paste0("GS.", names(traitOfInterest))
        names(GSPvalue) <- paste0("p.GS.", names(traitOfInterest))
        write.csv(merge(geneTraitSignificance,  GSPvalue, by=0, all = TRUE, sort = FALSE), 
                   paste0(SAVEPATH, colnames(phenWGCNA)[ j],"_GeneTrait_Significance_GSPvalue_",
                           Setting, "CutHeight",  CutHeight[i],
                           ".csv"), col.names = TRUE, row.names = FALSE)
      }
    }
  }
}

# 4. Coexpressed Pairs 
mainDir = "outputs/WGCNA/"
subDir = "CoexpressedPairs"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(length(list.files(recursive = TRUE, pattern = "Power.*.RData", path = paste0(mainDir, subDir))) == 0) {
  filenames <- list.files(recursive = TRUE, pattern = "All_genes_modules_Power", 
                            path = "outputs/WGCNA/MergeModules/", full.names = TRUE)	
  SAVEPATH <- "outputs/WGCNA/CoexpressedPairs/"
  for(file in filenames){
    args = file
    Genes_Modules <- read.csv(args[1])
    # remove genes which were not clustered(grey module)
    Genes_Modules <- Genes_Modules[Genes_Modules$mergedColors != "grey",]
    Modules <- unique(Genes_Modules$mergedColors)
    All_pairs <- c()
    for(module in 1 : length(Modules)) { 
      subset <- Genes_Modules[Genes_Modules$mergedColors == Modules[module],]
      # subset$NAMES <- gsub("\\..*", "", subset$NAMES)
      Genes <- as.character(sort(subset$NAMES))
      Pairs <- as.data.frame(t(combn(Genes, 2,)))
      # Pairs$module <- Modules[module] 
      All_pairs [[ module]] <- Pairs
      }
    All_pairs <- as.data.frame(do.call(rbind, All_pairs))
    print(head(All_pairs))
    print(dim(All_pairs))
    # All_pairs$module <- NULL
    save(All_pairs, file = paste0(SAVEPATH,
                                    gsub(".*All_genes_modules_|.csv", "", args[1]), ".RData"))
  }
}

# 5. Reactome knowledge pairs
mainDir = "outputs/WGCNA/"
subDir = "KnowledgeNetwork"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "Reactome_KnowledgePairs.RData"))) {
  Gene_list <- unique(rownames(v_BatchMus))

  # Annotating gene ids using gProfileR
  GenePathway_all <- c()
  for(i in seq(0, length(Gene_list), 100)) {
    print(i)
    GS <- as.vector(na.omit(Gene_list [(i + 1) :(i + 100)]))
    BG <- unique(as.vector(Gene_list))
    print(length(GS))
    GP <- gProfileR::gprofiler(as.vector(GS), organism = 'hsapiens', significant = F,
                                domain_size = "known", custom_bg = BG, correction_method = "fdr", src_filter = "REAC")
    if(dim(GP)[1] != 0) { GP$query.number <- i }
    if(i == 0) GenePathway_all <- GP else GenePathway_all <- rbind(GenePathway_all, GP)
  }
  save(GenePathway_all, file = "outputs/WGCNA/KnowledgeNetwork/GenePathway.RData")

  load("outputs/WGCNA/KnowledgeNetwork/GenePathway.RData")
  GenePathway_all [, c(1 : 3, 5, 7, 8, 11, 13)] <- NULL

  GenePathway_all <- aggregate(GenePathway_all[- c(1, 3, 4, 5)], by = list(GenePathway_all$term.size,
                                                                                 GenePathway_all$term.id,
                                                                                 GenePathway_all$ domain,
                                                                                 GenePathway_all$term.name),
                                c)
  colnames(GenePathway_all)[1 : 4] <- c("term.size", "term.id", "domain", "term.name")
  x <- GenePathway_all$overlap.size
  y <- character(length(x))
  for(i in 1 : length(x)) y[i] <- as.numeric(sum(x [[ i]]))
  GenePathway_all$overlap.size <- as.numeric(y)
  x <- GenePathway_all$intersection
  y <- character(length(x)) 
  for(i in 1 : length(x)) y[i] <- paste0(unique(x [[ i]]), collapse = ",")
  GenePathway_all$intersection <- y

  # Extracting the annotations from Reactome and HPO
  GenePathway_REAC <- GenePathway_all[grep("REAC:", GenePathway_all$term.id),]
  # Removing generic terms
  GenePathway_REAC <- GenePathway_REAC[! GenePathway_REAC$term.name == "Reactome",]

  rm(list = setdiff(ls(), c("GenePathway_REAC")))

  # Knowledge pairs based on Reactome and HP
  Pathways <- unique(GenePathway_REAC$term.id)
  All_pairs <- c()
  for(i in 1 : length(Pathways)) {
    genes <- sort(unlist(strsplit(GenePathway_REAC[GenePathway_REAC$term.id == Pathways[i], "intersection"], split = ",")))
    if(length(genes) != 1)
      { Pairs <- as.data.frame(t(combn(sort(genes), 2)))
      if(length(All_pairs) == 0) All_pairs <- Pairs else All_pairs <- rbind(All_pairs, Pairs)
    }
  }
  All_Knowledge_pairs <- unique(All_pairs)
  Annotated_genes <- unique(unlist(strsplit(GenePathway_REAC$intersection, split = ","))) 

  save(All_Knowledge_pairs, Annotated_genes, GenePathway_REAC, file = "outputs/WGCNA/KnowledgeNetwork/Reactome_KnowledgePairs.RData")
}

# 6. Overlap
mainDir = "outputs/WGCNA/"
subDir = "OverlabCP_KP"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "OverlapCP_KP.RData"))) {
  rm(list = ls())
  load("outputs/WGCNA/KnowledgeNetwork/Reactome_KnowledgePairs.RData")
  All_Knowledge_pairs <- sapply(All_Knowledge_pairs, as.character)
  filenames <- list.files(recursive = TRUE, path = "outputs/WGCNA/CoexpressedPairs", full.names = TRUE)	

  # length(Annotated_genes) ## No. of genes with Reatome annotation in GSE36398 dataset
  Overlap_all <- c()
  for(file in filenames){
    load(file)
    Coexpressed_pairs <- sapply(All_pairs, as.character) ; rm(All_pairs)
    Annotated_Coexpressed_pairs <- unique(Coexpressed_pairs[Coexpressed_pairs [,"V1"] %in% 
                                                                 Annotated_genes &
                                                                 Coexpressed_pairs [,"V2"] %in%
                                                                 Annotated_genes,])
    Annotated_Coexpressed_genes <- unique(c(Annotated_Coexpressed_pairs [,"V1"],
                                              Annotated_Coexpressed_pairs [,"V2"]))
    N <- length(Annotated_Coexpressed_genes) ## Co-expressed and annotated genes
    All_possible_pairs <-(length(Annotated_genes) *(length(Annotated_genes) - 1)) / 2 

    # Annotated_Coexpressed_pairs <- apply(Annotated_Coexpressed_pairs[, c("V1", "V2")], 1, FUN = function(x) paste(sort(x), collapse = "-"))
    Pairs <- merge(Annotated_Coexpressed_pairs, All_Knowledge_pairs)
    print(dim(Pairs))
    Annotated_Coexpressed_pairs <- Annotated_Coexpressed_pairs [, c(2, 1)]
    colnames(Annotated_Coexpressed_pairs) <- colnames(All_Knowledge_pairs)
    Pairs1 <- merge(Annotated_Coexpressed_pairs, All_Knowledge_pairs)
    print(dim(Pairs1))
    Pairs <- unique(rbind(Pairs, Pairs1))
    print(dim(Pairs))
    SM_SP <- dim(Pairs)[1]
    SM_nSP <- dim(Annotated_Coexpressed_pairs)[1] - SM_SP
    nSM_SP <- dim(All_Knowledge_pairs)[1] - SM_SP
    nSM_nSP <- All_possible_pairs -(SM_SP + SM_nSP + nSM_SP)
    table2_2 <- rbind(c(SM_SP, nSM_SP), c(SM_nSP, nSM_nSP))
    Ftest <- fisher.test(table2_2)
    OutPut <- data.frame(setting = gsub(".*\\/|.RData", "", file),
                          No_Annotated_Coexpressed_genes = length(Annotated_Coexpressed_genes),
                          Ftest = Ftest$p.value, odds_ratio = Ftest$estimate,
                          SM_SP = SM_SP, SM_nSP = SM_nSP, nSM_SP = nSM_SP, nSM_nSP = nSM_nSP)
    rownames(OutPut) <- NULL
    print(OutPut)
    Overlap_all [[gsub(".*\\/|.RData", "", file)]] <- OutPut
  }
  Overlap_all <- data.table::rbindlist(Overlap_all)
  save(Overlap_all, file = "outputs/WGCNA/OverlabCP_KP/OverlapCP_KP.RData")
}

# 7. Settings summary
if(! file.exists("outputs/WGCNA/OverlabCP_KP/SettingsSummary.RData")) {
  rm(list = ls())
  load("outputs/WGCNA/OverlabCP_KP/OverlapCP_KP.RData")
  # head(Overlap_all)
  filenames <- list.files(recursive = TRUE, path = "outputs/WGCNA/MergeModules/", pattern = "All_genes_modules", full.names = TRUE)
  SettingsSummary <- matrix(NA, nrow = length(filenames), ncol = 8)
  rownames(SettingsSummary) <- gsub(".*All_genes_modules_|.csv", "", filenames)
  for(file in filenames) {
    x <- read.csv(file)
    setting <- gsub(".*All_genes_modules_|.csv", "", file)
    SettingsSummary[setting,] <- c(length(unique(x[x$mergedColors == "grey", 1])), 
                                       length(table(x$mergedColors)) - 1,
                                       summary(as.vector(table(x [x$mergedColors != "grey",]$mergedColors))))
    }
  colnames(SettingsSummary) <- c("GreyM.size", "NoModules", "MinM.Size", "1stQ", "MedianM.Size",
                                   "MeanM.Size", "3stQ", "MaxM.Size")
  # The final setting table
  Setting <- merge(Overlap_all, as.data.frame(SettingsSummary) %>% 
                      tibble::rownames_to_column(var = "setting"), by = "setting")
  save(Setting, file = "outputs/WGCNA/OverlabCP_KP/SettingsSummary.RData") } else {
    rm(list = prob::setdiff(ls(), c("v_BatchMus", "Data")))
    load("outputs/WGCNA/OverlabCP_KP/SettingsSummary.RData")
  }
```

### Selecting the best WGCNA setting
```{r SelectingBestSetting, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
Setting <- Setting %>% 
  mutate(Power = as.numeric(gsub("Power|MinModuleSize.*", "", setting))) %>%
  mutate(MinModuleSize =  gsub(".*MinModuleSize|deepSplit.*", "", setting)) %>%
  mutate(deepSplit =  gsub(".*deepSplit|CutHeight.*", "", setting)) %>%
  mutate(CutHeight =  gsub(".*CutHeight", "", setting)) %>%
  arrange(Power)
p1 <- ggplot(Setting) +
  stat_summary(mapping = aes(x = reorder(Power, setting), y = odds_ratio),
                geom = "point", fun = "median") +
  labs(x = "Power", y = "Reactome's enrichment factor") +
  theme_bw() +
  theme(axis.text = element_text(size = 10, colour = "black"),
         axis.title = element_text(size= 10, face = "bold"), panel.grid = element_blank())

p2 <- ggplot(Setting) +
  stat_summary(mapping = aes(x = MinModuleSize, y = odds_ratio),
                geom = "point", fun = "median") +
  labs(x = "minClusterSize") +
  theme_bw() +
  theme(axis.text = element_text(size = 10, colour = "black"),
         axis.title = element_text(size= 10, face = "bold"), 
         axis.title.y = element_blank(), panel.grid = element_blank())

p3 <- ggplot(Setting) +
  stat_summary(mapping = aes(x = deepSplit, y = odds_ratio),
                geom = "point", fun = "median") +
  labs(x = "deepSplit") +
  theme_bw() +
  theme(axis.text = element_text(size = 10, colour = "black"),
         axis.title = element_text(size= 10, face = "bold"), 
         axis.title.y = element_blank(), panel.grid = element_blank())

p4 <- ggplot(Setting) +
  stat_summary(mapping = aes(x = CutHeight, y = odds_ratio),
                geom = "point", fun = "median") +
  labs(x = "cutHeight") +
  theme_bw() +
  theme(axis.text = element_text(size = 10, colour = "black"),
         axis.title = element_text(size= 10, face = "bold"), 
         axis.title.y = element_blank(), panel.grid = element_blank())

grid.newpage()
grid.draw(cbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p3),  
                    ggplotGrob(p4),                 
                    size = "first")) 
rm(list = paste0("p", 1:4))
ggplot(Setting, aes(x = NoModules, y = odds_ratio)) +
  geom_point(size = 2, alpha = 0.5) +
  stat_smooth(method = "lm", col = "red") +
  theme_bw() +
  labs(x = "Number of modules", y = "Reactome's enrichment factor") +
  theme(axis.text = element_text(size = 10, colour = "black"),
          axis.title = element_text(size= 10, face = "bold"), 
          title = element_text(size= 10, colour = "black"), panel.grid = element_blank())

# Selecting the best setting considering the grey module size
# < %30 in grey module
print("< %30 in grey module")
SettingSubset0.3 <- Setting[Setting$GreyM.size / nrow(v_BatchMus) < 0.3,]
# Selected setting
print(SettingSubset0.3[which.max(SettingSubset0.3$odds),])

ggplot(SettingSubset0.3, aes(x = NoModules, y = odds_ratio)) +
  geom_point(size = 2, alpha = 0.5) +
  stat_smooth(method = "lm", col = "red") +
  theme_bw() +
  labs(x = "Number of modules", y = "Reactome's enrichment factor") +
  theme(axis.text = element_text(size = 10, colour = "black"),
          axis.title = element_text(size= 10, face = "bold"), 
          title = element_text(size= 10, colour = "black"), panel.grid = element_blank())


# < %50 in grey module
print("< %50 in grey module")
SettingSubset0.5 <- Setting[Setting$GreyM.size / nrow(v_BatchMus) < 0.5,]
# Selected setting
print(SettingSubset0.5[which.max(SettingSubset0.5$odds),])

ggplot(SettingSubset0.5, aes(x = NoModules, y = odds_ratio)) +
  geom_point(size = 2, alpha = 0.5) +
  stat_smooth(method = "lm", col = "red") +
  theme_bw() +
  labs(x = "Number of modules", y = "Reactome's enrichment factor") +
  theme(axis.text = element_text(size = 10, colour = "black"),
          axis.title = element_text(size= 10, face = "bold"), 
          title = element_text(size= 10, colour = "black"), panel.grid = element_blank())
```

The optimal set of parameters with the highest enrichment factor was: power: 8, MinModuleSize: 20, deepSplit: 0, Cut Height: 0.2.

```{r}
sessionInfo()
```
