---
title: "Consensus WGCNA"
author:
- tabbassidaloii
output:
 md_document:
  variant: markdown_github
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

'Note:' The muscle abbreviations were GRA(for Gracilis), SED(semitendinosus distal),  SEM(semitendinosus middle)), GAL(gastrocnemius lateralis), REF(rectus femoris), VAM(vastus medialis) and VAL(vastus lateralis) in all the analyses(scripts) and they were only changed to GR, STM, STD, GL, RF, VM and VL for the plots(shown in the publication).

```{r setup, include = FALSE}
rm(list = ls())
if(!"limma" %in% installed.packages())BiocManager::install("limma") 
if(!"edgeR" %in% installed.packages())BiocManager::install("edgeR") 
if(!"WGCNA" %in% installed.packages())BiocManager::install("WGCNA") 
if(!"dplyr" %in% installed.packages())install.packages("dplyr")
if(!"ggplot2" %in% installed.packages())install.packages("ggplot2")

suppressPackageStartupMessages({ 
  library(WGCNA)
  library(limma)
  library(edgeR)
  library(dplyr)
  library(ggplot2)
 })
opts_chunk$set(eval = TRUE, echo = TRUE, tidy = TRUE, highlight = TRUE, dev = c('png'), fig.width = 10, fig.height = 8,  fig.path = "figures/")

BiocStyle::markdown()
options(scipen = 999)

# Loading the data
Data <- new.env(); load("outputs/correctedGenesDatasetToAnalysis.RData", envir = Data)
D_BatchMus <- Data$D_BatchMus
# Creating a design matrix
design <- model.matrix(~ group + individual, data = D_BatchMus$samples)
colnames(design) <- gsub("group|individual", "", colnames(design))
# mean-variance trend was removed using the voom
v_BatchMus <- voom(D_BatchMus, design, plot = TRUE)
```

## Gene coexpression network analysis

### Consensus WGCNA
##### Candidate setting based on overlap with knowledge network
##### Power = 8;  MinModuleSize = 20; deepSplit = 0; CutHeight = 0.2
we decided to make the consensus network on TOM level because 
-(1) the adjacency is calculated only based on 5 to 7 data points 
-(2) the TOM includes the topological overlaps of genes which add more information to calculate the consensus matrices

1. Creating adjacency matrices for each individual
2. Calculating TOM 
3. Scaling the TOM
4. Consensus Network(quantile 0.1, 0.25, 0.3, 0.5, 0.6, 0.75, and 0.8)
5. Calculating dissTOM
6. Detecting modules
7. Merging modules
8. Defining co-expressed pairs(CPs)
9. Finding the overlap between KPs and CPs
10. Creating summary file showing the overlap of each network with knowledge network

```{r AllStepsWGCNA, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
enableWGCNAThreads()
# Candidate setting based on overlap with knowledge network
# Power = 8;  MinModuleSize = 20; deepSplit = 0; CutHeight = 0.2

# Steps 1 to 5(on a cluster)
SHARK = FALSE # The files are generated 
mainDir = "outputs/consensusWGCNA/"
subDir = "MergeModules"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "All_genes_modules_Consensus0.1.csv"))){
  subDir = "dissTOM"
  if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
  if(SHARK){
    # 1. Adjacency matrix for each individual
    nSets <- nlevels(v_BatchMus$targets$individual)
    nGenes <- nrow(v_BatchMus$E)
    # Initialize an appropriate array to hold the expression set for each individual
    SampleSets <- lapply(split(v_BatchMus$targets, v_BatchMus$targets$individual),
                         function(x) rownames(x))
    multiExpr <- lapply(setNames(names(SampleSets), names(SampleSets)),
                        function(x) v_BatchMus$E[, grep(x, colnames(v_BatchMus$E))])
    # Create correlation matrix
    # Initialize an appropriate array to hold the adjacencies
    Adjacencies <- array(0, dim = c(nSets, nGenes, nGenes));
    # Calculate adjacencies in each individual data set
    for(set in 1:nSets){
      Adjacencies[set, ,] <- adjacency(t(multiExpr[[set]]), power = 8, 
                                         type = "signed hybrid", corFnc = "bicor", 
                                         corOptions = c("maxPOutliers = 0.1"))
      }
    
    # 2. TOM
    # Define a similarity based on the topological overlap
    # Initialize an appropriate array to hold the TOMs
    TOM <- array(0, dim = c(nSets, nGenes, nGenes))
    # Calculate TOMs in each individual data set
    for(set in 1:nSets){
      TOM[set, ,] <- WGCNA::TOMsimilarity(Adjacencies[set, ,], TOMType = "signed")
      }
    
    # 3. Scaling the TOM
    # Define the reference percentile
    scaleP = 0.95
    TOMScalingSamples <- list()
    # These are TOM values at reference percentile
    scaleQuant <- rep(1, nSets)
    # Scaling powers to equalize reference TOM values
    scalePowers <- rep(1, nSets)
    # Loop over sets
    for(set in 1:nSets){
      # Select the sampled TOM entries
      TOMScalingSamples[[set]] <- as.dist(TOM[set, ,])
      # Calculate the 95th percentile
      scaleQuant[set] <- quantile(TOMScalingSamples[[set]],
                                    probs = scaleP, type = 8);
      # Scale the male TOM
      if(set>1){
        scalePowers[set] = log(scaleQuant[1])/log(scaleQuant[set]);
        TOM[set, ,] = TOM[set, ,]^scalePowers[set];
        }
      }
      
    # 4. Consensus Network
    for(pscale in c(0.1, 0.2, 0.25, 0.3, 0.4, 0.5, 0.6, 0.7, 0.75, 0.8)){
        consensusTOM <- pquantile(prob = pscale, 
                                  TOM[1, ,], TOM[2, ,], TOM[3, ,], TOM[4, ,], TOM[5, ,],
                                  TOM[6, ,], TOM[7, ,], TOM[8, ,], TOM[9, ,], TOM[10, ,], 
                                  TOM[11, ,], TOM[12, ,], TOM[13, ,], TOM[14, ,], TOM[15, ,], 
                                  TOM[16, ,], TOM[17, ,], TOM[18, ,], TOM[19, ,], TOM[20, ,])
    
        colnames(consensusTOM) <- rownames(consensusTOM) <- rownames(v_BatchMus$E)
        # 5. dissTOM
        dissTOM <- 1 - consensusTOM
        save(dissTOM, file = paste0("outputs/consensusWGCNA/dissTOM/dissTOM_Consensus", pscale, ".RData"))
      }
    }
  
  filenames <- list.files(path = "outputs/consensusWGCNA/dissTOM/", full.names = T)
  
  for(dissTOMfile in filenames){
    load(dissTOMfile)
    pscale <- gsub(".*Consensus|.RData", "", dissTOMfile)
    # Hierarchical gene clustering(according to the dissimilarity matrix)
    geneTree <- flashClust::flashClust(as.dist(dissTOM), method = "average")
    save(dissTOM, geneTree, file = paste0("outputs/consensusWGCNA/dissTOM/dissTOM_Consensus", pscale, ".RData"))
    # 6. Module Detection
    # Power = 8;  MinModuleSize = 20; deepSplit = 0; CutHeight = 0.2
    # Module detection
    # Set the minimum module size 
    minModuleSize <- 20
    deepSplit <- 0
    # Module detection by cutting branches
    dynamicMods <- dynamicTreeCut::cutreeDynamic(dendro = geneTree, distM = dissTOM,
                                                 method = "hybrid", deepSplit = deepSplit,
                                                 pamRespectsDendro = FALSE,
                                                 minClusterSize = minModuleSize)
    # Convert labels to colors for plotting
    dynamicColors <- WGCNA::labels2colors(dynamicMods)
    # Calculate eigengenes(clustring modules based on expression similarities)
    MEList <- WGCNA::moduleEigengenes(t(v_BatchMus$E), colors = dynamicColors)
    MEs <- MEList$eigengenes
    NAMES <- colnames(dissTOM)
    save(dynamicColors, geneTree, NAMES, file = paste0("outputs/consensusWGCNA/ModuleDetection/ModuleDetection_Consensus", pscale, ".RData"))
    # 7. Merge Modules
    SAVEPATH <- "outputs/consensusWGCNA/MergeModules/"
    # Merging similar modules
    CutHeight <- 0.2
    # Call an automatic merging function
    merge <- WGCNA::mergeCloseModules(t(v_BatchMus$E), dynamicColors,
                                       cutHeight = CutHeight,
                                       verbose = 3, corFnc = bicor) 
    # The merged module colors
    mergedColors <- merge$colors
    # Eigengenes of the new merged modules:
    mergedMEs <- merge$newMEs
    write.csv(mergedMEs, paste0(SAVEPATH, "ModuleEigengenes_merged_Consensus", pscale, ".csv"),
               row.names = FALSE)
    # Summary output of network analysis results(after merging)
    module_colors <- unique(mergedColors)
    All_genes_modules <- as.data.frame(cbind(NAMES, mergedColors)) 
    # colnames(All_genes_modules) <- NULL
    write.csv(All_genes_modules, paste0(SAVEPATH, "All_genes_modules_Consensus", 
                                        pscale, ".csv"), row.names = FALSE)
    # names(colors) of the modules
    nSamples <- nrow(t(v_BatchMus$E))
    modNames <- substring(names(mergedMEs), 3)
    geneModuleMembership <- as.data.frame(bicor(t(v_BatchMus$E), mergedMEs, 
                                                maxPOutliers = 0.1)) 
    MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples)) 
    names(geneModuleMembership) <- gsub("ME", "MM", names(geneModuleMembership))
    names(MMPvalue) <-  gsub("ME", "p.MM", names(MMPvalue))
    write.csv(merge(geneModuleMembership, MMPvalue, by=0, all = TRUE, sort = FALSE),
              paste0(SAVEPATH, "GeneModule_Membership_MMPvalue_Consensus", pscale, ".csv"),
              col.names = TRUE, row.names = FALSE)
    phenWGCNA <- v_BatchMus$targets %>%
      mutate(group = as.numeric(group)) %>%
      mutate(individual = as.numeric(gsub("MD", "", individual))) %>%
      select(-"sampleID") %>%
      mutate(IsolationRNA = ifelse(IsolationRNA == "WithOutKit", 1, 0)) %>%
      mutate(Hospital = ifelse(Hospital == "HMC", 1, 0)) %>%
      mutate(lane = as.numeric(lane)) %>%
      mutate(batch = as.numeric(batch))
    # Module membership values and gene significance
    for(j in 1 : length(phenWGCNA)[1]){
      # Define interested variable of datTrait
      traitOfInterest <- as.data.frame(phenWGCNA[, j])
      names(traitOfInterest) <-  colnames(phenWGCNA)[j]
      ## Relating modules to external clinical traits and identifying important genes
      nGenes <- ncol(t(v_BatchMus$E))
      nSamples <- nrow(t(v_BatchMus$E))
      moduleTraitCor <- matrix(nrow = dim(mergedMEs)[2], ncol = dim(phenWGCNA)[2]) 
      colnames(moduleTraitCor) <- paste0("Cor_", colnames(phenWGCNA))
      rownames(moduleTraitCor) <- colnames(mergedMEs) 
      for(trait in colnames(phenWGCNA)){
        moduleTraitCor[, grep(trait, colnames(moduleTraitCor))] <-
            cor(mergedMEs, phenWGCNA[, trait], use = "pairwise.complete.obs")
        }
      
      moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples)
      colnames(moduleTraitPvalue) <- paste0("Pvalue_", colnames(moduleTraitPvalue))
      # Writing the tables down with modules and correlation
      write.csv(merge(moduleTraitCor, moduleTraitPvalue, by = 0, all = TRUE, 
                      sort = FALSE),
                 paste0(SAVEPATH, "ModuleTrait_Cor_Pvalue_Consensus", pscale, ".csv"), 
                 col.names = TRUE, row.names = FALSE) 
      # correlation matrix of each gene and each traits
      geneTraitSignificance <- as.data.frame(cor(t(v_BatchMus$E), traitOfInterest,
                                                   use = 'pairwise.complete.obs'))
      GSPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))
      names(geneTraitSignificance) <- paste0("GS.", names(traitOfInterest))
      names(GSPvalue) <- paste0("p.GS.", names(traitOfInterest))
      write.csv(merge(geneTraitSignificance,  GSPvalue, by=0, all = TRUE, sort = FALSE), 
                paste0(SAVEPATH, colnames(phenWGCNA)[j], "_GeneTrait_Significance_GSPvalue_Consensus", pscale, ".csv"), 
                col.names = TRUE, row.names = FALSE)
      }
    }
  }

# 8. Coexpressed Pairs 
mainDir = "outputs/consensusWGCNA/"
subDir = "CoexpressedPairs"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "/Consensus0.1.RData"))){
  filenames <- list.files(recursive = TRUE, pattern = "All_genes_modules", 
                          path = "outputs/consensusWGCNA/MergeModules/", full.names = TRUE)	
  SAVEPATH <- "outputs/consensusWGCNA/CoexpressedPairs/"
  for(file in filenames){
    Genes_Modules <- read.csv(file)
    # remove genes which were not clustered(grey module)
    Genes_Modules <- Genes_Modules[Genes_Modules$mergedColors != "grey",]
    Modules <- unique(Genes_Modules$mergedColors)
    All_pairs <- c()
    for(module in 1 : length(Modules)){ 
      subset <- Genes_Modules[Genes_Modules$mergedColors == Modules[module],]
      # subset$NAMES <- gsub("\\..*", "", subset$NAMES)
      Genes <- as.character(sort(subset$NAMES))
      Pairs <- as.data.frame(t(combn(Genes, 2,)))
      # Pairs$module <- Modules[module] 
      All_pairs[[module]] <- Pairs
      }
    All_pairs <- as.data.frame(do.call(rbind, All_pairs))
    print(head(All_pairs))
    print(dim(All_pairs))
    # All_pairs$module <- NULL
    save(All_pairs, file = paste0(SAVEPATH,
                                  gsub(".*All_genes_modules_|.csv", "", file), ".RData"))
    }
  }

# 9. Overlap
mainDir = "outputs/consensusWGCNA/"
subDir = "OverlabCP_KP"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "/OverlapCP_KP.RData"))){
  load("outputs/WGCNA/KnowledgeNetwork/Reactome_KnowledgePairs.RData")
  
  All_Knowledge_pairs <- sapply(All_Knowledge_pairs, as.character)
  filenames <- list.files(recursive = TRUE, path = "outputs/consensusWGCNA/CoexpressedPairs", full.names = TRUE)	
  # length(Annotated_genes) ## No. of genes with Reatome annotation in GSE36398 dataset
  Overlap_all <- c()
  for(file in filenames){
    load(file)
    Coexpressed_pairs <- sapply(All_pairs, as.character) ; rm(All_pairs)
    Annotated_Coexpressed_pairs <- unique(Coexpressed_pairs[Coexpressed_pairs[, "V1"] %in% 
                                                                 Annotated_genes &
                                                                 Coexpressed_pairs[, "V2"] %in%
                                                                 Annotated_genes,])
    Annotated_Coexpressed_genes <- unique(c(Annotated_Coexpressed_pairs[, "V1"],
                                            Annotated_Coexpressed_pairs[, "V2"]))
    N <- length(Annotated_Coexpressed_genes) ## Co-expressed and annotated genes
    All_possible_pairs <-(length(Annotated_genes) *(length(Annotated_genes) - 1)) / 2 

    # Annotated_Coexpressed_pairs <- apply(Annotated_Coexpressed_pairs[, c("V1", "V2")], 1, FUN = function(x) paste(sort(x), collapse = "-"))
    Pairs <- merge(Annotated_Coexpressed_pairs, All_Knowledge_pairs)
    print(dim(Pairs))
    # Annotated_Coexpressed_pairs <- Annotated_Coexpressed_pairs[, c(2, 1)]
    # colnames(Annotated_Coexpressed_pairs) <- colnames(All_Knowledge_pairs)
    # Pairs1 <- merge(Annotated_Coexpressed_pairs, All_Knowledge_pairs)
    # print(dim(Pairs1))
    # Pairs <- unique(rbind(Pairs, Pairs1))
    # print(dim(Pairs))
    SM_SP <- dim(Pairs)[1]
    SM_nSP <- dim(Annotated_Coexpressed_pairs)[1] - SM_SP
    nSM_SP <- dim(All_Knowledge_pairs)[1] - SM_SP
    nSM_nSP <- All_possible_pairs -(SM_SP + SM_nSP + nSM_SP)
    table2_2 <- rbind(c(SM_SP, nSM_SP), c(SM_nSP, nSM_nSP))
    Ftest <- fisher.test(table2_2)
    OutPut <- data.frame(Network = gsub(".*\\/|.RData", "", file),
                          No_Annotated_Coexpressed_genes = length(Annotated_Coexpressed_genes),
                          Ftest = Ftest$p.value, odds_ratio = Ftest$estimate,
                          SM_SP = SM_SP, SM_nSP = SM_nSP, nSM_SP = nSM_SP, nSM_nSP = nSM_nSP)
    rownames(OutPut) <- NULL
    print(OutPut)
    Overlap_all[[gsub(".*\\/|.RData", "", file)]] <- OutPut
    }
  Overlap_all <- data.table::rbindlist(Overlap_all)
  save(Overlap_all, file = "outputs/consensusWGCNA/OverlabCP_KP/OverlapCP_KP.RData")
  }

# 10. Settings summary
if(!file.exists(paste0(mainDir, subDir, "/", "SettingsSummary.RData"))){
  rm(list = ls())
  load("outputs/consensusWGCNA/OverlabCP_KP/OverlapCP_KP.RData")
  # head(Overlap_all)
  filenames <- list.files(recursive = TRUE, path = "outputs/consensusWGCNA/MergeModules/", pattern = "All_genes_modules", full.names = TRUE)
  SettingsSummary <- matrix(NA, nrow = length(filenames), ncol = 8)
  rownames(SettingsSummary) <- gsub(".*All_genes_modules_|.csv", "", filenames)
  for(file in filenames){
    x <- read.csv(file)
    setting <- gsub(".*All_genes_modules_|.csv", "", file)
    SettingsSummary[setting,] <- c(length(unique(x[x$mergedColors == "grey", 1])), 
                                   length(table(x$mergedColors)) - 1,
                                   summary(as.vector(table(x[x$mergedColors != "grey",]$mergedColors))))
    }
  colnames(SettingsSummary) <- c("GreyM.size", "NoModules", "MinM.Size", "1stQ", "MedianM.Size",
                                 "MeanM.Size", "3stQ", "MaxM.Size")
  # The final setting table
  Setting <- merge(Overlap_all, as.data.frame(SettingsSummary) %>% 
                      tibble::rownames_to_column(var = "Network"), by = "Network")
  save(Setting, file = "outputs/consensusWGCNA/OverlabCP_KP/SettingsSummary.RData")
    }else{
      rm(list = prob::setdiff(ls(), c("v_BatchMus", "Data")))
      load("outputs/consensusWGCNA/OverlabCP_KP/SettingsSummary.RData")
      }
 
```

### Selecting the proper quantile or percentile for the consensus network
```{r SelectingBestSetting, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
CorFun <- function(data1, data2){
  if(all(rownames(data1) != rownames(data2)) & 
      all(colnames(data1) != colnames(data2))) print("Check the input rownames")
  COR <- matrix(NA, nrow = ncol(data1), ncol = 1)
  rownames(COR) <- colnames(data1)
  CORPval <- COR
  for(i in colnames(data1)){
    COR[i, 1] <- cor(data1[, i], data2[, i])
    CORPval[i, 1] <- cor.test(data1[, i], data2[, i])$p.value
 }
  CorList <- list(Cor = COR, CorPval = CORPval)
  return(CorList)
}

# nonconsensus
load("outputs/WGCNA/dissTOM/dissTOMPower8.RData")
TOM <- 1 - dissTOM; rm(dissTOM)
All_genes_modules <- read.csv("outputs/WGCNA/MergeModules/All_genes_modules_Power8MinModuleSize20deepSplit0CutHeight0.2.csv")
all(colnames(TOM) == All_genes_modules$NAMES)
COL <- All_genes_modules$mergedColors
ConnectivityMat <- intramodularConnectivity(TOM, COL)

for(pscale in c(0.1, 0.2, 0.25, 0.3, 0.4, 0.5, 0.6, 0.7, 0.75, 0.8)){
  load(paste0("outputs/consensusWGCNA/dissTOM/dissTOM_Consensus", pscale, ".RData"))
  consensusTOM <- 1 - dissTOM; rm(dissTOM)
  Cor_NonconsensusConsensus <- CorFun(TOM, consensusTOM)
  colnames(Cor_NonconsensusConsensus$Cor) <- colnames(Cor_NonconsensusConsensus$CorPval) <- pscale
  assign(paste0("Cor_NonconsensusConsensus", pscale), Cor_NonconsensusConsensus)

  All_genes_modules <- read.csv(paste0("outputs/consensusWGCNA/MergeModules/All_genes_modules_Consensus", pscale, ".csv"))
  all(colnames(consensusTOM) == All_genes_modules$NAMES)
  COL <- All_genes_modules$mergedColors
  ConnectivityMat <- WGCNA::intramodularConnectivity(consensusTOM, COL)
  
  rm(consensusTOM)
  assign(paste0("ConnectivityMatConsensus", pscale), ConnectivityMat)
  # print(scaleFreeFitIndex(ConnectivityMat$kTotal, nBreaks = 10, removeFirst = FALSE))
  print(WGCNA::scaleFreePlot(ConnectivityMat$kTotal, nBreaks = 10, truncated = FALSE, removeFirst = FALSE, main = paste0("percentile: ", pscale, "\n")))
}

load("outputs/WGCNA/Adjacency/AdjacencyPower8.RData")
all(colnames(Adjacency) == All_genes_modules$NAMES)
ConnectivityMatAdj <- WGCNA::intramodularConnectivity(Adjacency, COL)
cor(ConnectivityMatAdj$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.1$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.2$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.25$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.3$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.4$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.5$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.6$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.7$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.75$kTotal, ConnectivityMat$kTotal)
cor(ConnectivityMatConsensus0.8$kTotal, ConnectivityMat$kTotal)

Consensus <- "Consensus0.4"
```

### Enrichment for celltype markers in modules
```{r MarkersInModules, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE, fig.height = 7, fig.width = 14}
filenames <- list.files(recursive = TRUE, path = "outputs/consensusWGCNA", pattern = paste0(Consensus, ".csv"), full.names = TRUE)

# Loading the data
Data <- new.env(); load("outputs/correctedGenesDatasetToAnalysis.RData", envir = Data)
D_BatchMus <- Data$D_BatchMus
# Creating a design matrix
design <- model.matrix(~ group + individual, data = D_BatchMus$samples)
colnames(design) <- gsub("group|individual", "", colnames(design))
# mean-variance trend was removed using the voom
v_BatchMus <- voom(D_BatchMus, design, plot = TRUE)

# Importing gene-modules table
GeneModules <- read.csv(filenames[grepl("All_genes_modules", filenames)]) %>%
  mutate(hgnc_symbol = v_BatchMus$genes$hgnc_symbol[match(NAMES, v_BatchMus$genes$ensembl_gene_id)])

# Rename modules(colors to numbers)
Module_rename <- data.frame(Module = prob::setdiff(sort(unique(GeneModules$mergedColors)), "grey"), 
                             Rename = paste("M", seq.int(length(unique(GeneModules$mergedColors))-1), sep = "."), stringsAsFactors = FALSE) 

GeneModules$mergedColors <- Module_rename$Rename[match(GeneModules$mergedColors, Module_rename$Module)] 
GeneModules$mergedColors[is.na(GeneModules$mergedColors)] = "grey"

# Importing marker list
load("inputs/cellTypeMarkers.RData")
CellTypes <- unique(markers$`Cell-type`)

# Markers distribution in the modules
CellMarkersInModules <- data.frame(table(GeneModules$mergedColors)) %>% mutate(Var1 = as.character(Var1))
colnames(CellMarkersInModules) <- c("Module", "Size")

for(cell in CellTypes){
  # cell markers
  GeneMarkers <- markers$Gene[markers$`Cell-type`== cell]
  
  # Modules in which the cell markers are 
  NoMarkers <- as.data.frame(table(GeneModules %>% filter(hgnc_symbol %in% GeneMarkers) %>% .$mergedColors))
  colnames(NoMarkers) <- c("Module", gsub(" ", "", cell))
  
  CellMarkersInModules <- merge(CellMarkersInModules, NoMarkers, all = T)
}
CellMarkersInModules[is.na(CellMarkersInModules)] = 0

CellMarkersInModulesPlot <- CellMarkersInModules %>% 
  dplyr::rename(`Endothelial Cells` = EndothelialCells, `LUM+ FAP Cells` = `LUM+FAPCells`,
          `Satellite Cells` = SatelliteCells, `FBN1+ FAP Cells` = `FBN1+FAPCells`,
          `PCV Endothelial Cells` = PCVEndothelialCells, `T and B Cells` = TandBCells,
          `Smooth Muscle Cells` = SmoothMuscleCells, `Myeloid Cells` = MyeloidCells,
          `Fast twitch` = Fasttwitch, `Slow twitch` = Slowtwitch, `NK Cells` = NKCells) %>% 
  mutate(Module = paste0(Module, "(", Size, " genes)")) %>% 
  reshape2::melt(-c(1, 2), id.var = "Module") %>%
  filter(Module != "grey(1582 genes)")

CellMarkersInModulesPlot <- CellMarkersInModulesPlot[order(as.numeric(gsub("M.| .*", "", CellMarkersInModulesPlot$Module))),] %>% mutate(Module = factor(Module, levels = unique(Module)))

 
ORDER <- c("M.14(162 genes)" = 1, "M.16(248 genes)" = 2, "M.23(1125 genes)" = 3, "M.4(262 genes)" = 4, 
           "M.33(538 genes)" = 5, "M.31(945 genes)" = 6, "M.18(220 genes)" = 7,"M.13(300 genes)" = 8,
           "M.24(587 genes)" = 9, "M.17(176 genes)" = 10, "M.7(308 genes)" = 11, "M.3(139 genes)" = 12, 
           "M.32(236 genes)" = 13, "M.30(75 genes)" = 14, "M.15(169 genes)" = 15, "M.12(429 genes)" = 16,
           "M.6(1459 genes)" = 17, "M.10(1376 genes)" = 18, "M.11(299 genes)" = 19, "M.25(188 genes)" = 20,
           "M.5(190 genes)" = 21, "M.8(127 genes)" = 22, "M.27(222 genes)" = 23, "M.34(628 genes)" = 24,
           "M.29(38 genes)" = 25, "M.21(113 genes)" = 26, "M.1(136 genes)" = 27, "M.26(228 genes)" = 28,
           "M.9(732 genes)" = 29, "M.19(825 genes)" = 30, "M.22(368 genes)" = 31, "M.28(333 genes)" = 32,
           "M.20(181 genes)" = 33, "M.2(625 genes)" = 34, "grey(1582 genes)" = 35)

      
CellMarkersInModulesPlot$order <- ORDER[match(CellMarkersInModulesPlot$Module, names(ORDER))]
# Add hyphen 
levels(CellMarkersInModulesPlot$variable)[levels(CellMarkersInModulesPlot$variable) == "Slow twitch"] = "Slow-twitch"
levels(CellMarkersInModulesPlot$variable)[levels(CellMarkersInModulesPlot$variable) == "Fast twitch"] = "Fast-twitch"

(Tosave <- ggplot(CellMarkersInModulesPlot, aes(x = forcats::fct_reorder(Module, order), y = variable)) + 
  geom_tile(aes(fill = value), width = 0.9, height = 0.9) + 
  scale_fill_gradientn(colours = c("white", "red"),
                        values = scales::rescale(c(0, 15, 31))) +
  geom_text(aes(label = value)) + 
  # ggtitle("Loading of cell type markers in modules")  +
  labs(x = NULL, y = NULL, fill = "No. Genes") +
  theme(axis.text = element_text(size = 12), axis.ticks = element_blank(),
         axis.text.x = element_text(angle = 90, hjust =  1, vjust = 0.5),
         axis.title = element_text(size= 12, face = "bold"), legend.position = "none") +
  scale_y_discrete(position = "right"))
ggsave(Tosave, device = "tiff", units = "cm", width = 18, height = 12, filename = "figures/Modules_cellMArkers.tif")
```

### Module-trait association
```{r ModuleTrait, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 10}
# Sample information
phenWGCNA <- v_BatchMus$targets[, c(1, 5, 12)]

# Module eigengene(ME) vector
mergedMEs <- read.csv(filenames[grepl("ModuleEigengenes", filenames)]) 
rownames(mergedMEs) <- colnames(v_BatchMus)
colnames(mergedMEs)[colnames(mergedMEs) != "MEgrey"] <- Module_rename$Rename[match(gsub("ME", "", colnames(mergedMEs)[colnames(mergedMEs) != "MEgrey"]), Module_rename$Module)] 
colnames(mergedMEs)[colnames(mergedMEs) == "MEgrey"] = "grey"

# Gene modules
Module_sizes <- as.data.frame(table(GeneModules$mergedColors))

# Final table with sample informations and module eigengene 
complete_info <- merge(phenWGCNA, mergedMEs, by = "row.names") 

Module_list <- colnames(complete_info)[grep("M.", colnames(complete_info))]

ModuleTrait <- matrix(NA, nrow = length(Module_list), ncol = 2)
rownames(ModuleTrait) <- Module_list
colnames(ModuleTrait) <- c("Muscle", "Individual")
library(lmerTest)
for(module in Module_list){
  fml <- as.formula(paste(noquote(module), "~ group +(1|individual)"))
  model <- lmerTest::lmer(fml, data = complete_info, REML = TRUE)
  ModuleTrait[module,] <- c(anova(model)$`Pr(>F)`, ranova(model)$`Pr(>Chisq)`[2])
}

ModuleTraitAdj <- ModuleTrait
for(factor in colnames(ModuleTrait)){
  ModuleTraitAdj[, factor] <- round(p.adjust(ModuleTrait[, factor], method = "fdr"), 6)
}
ModuleTraitAdj <- as.data.frame(ModuleTraitAdj, as.is = T)
ModuleTraitAdj$Module <- paste0(Module_sizes$Var1, "(", Module_sizes$Freq, ")")[match(noquote(rownames(ModuleTraitAdj)), Module_sizes$Var1)]

ModuleTraitAdjPlot <- reshape2::melt(ModuleTraitAdj) 
ModuleTraitAdjPlot <- ModuleTraitAdjPlot[order(as.numeric(gsub("M.| .*", "", ModuleTraitAdjPlot$Module))),] %>% mutate(Module = factor(Module, levels = unique(Module))) 

ggplot(ModuleTraitAdjPlot, aes(x = variable, y = Module)) + 
  geom_tile(aes(fill = value)) + 
  scale_fill_gradientn(colours = c("white", "red", "black"),
                        values = scales::rescale(c(0, 0.05, 0.1, 0.5, 1))) +
  geom_text(aes(label = format(value, digits = 2))) +
  ggtitle("Modules-factors association")  +
  labs(x = NULL, y = NULL, fill = "FDR") +
  theme(axis.text = element_text(size = 16),
         axis.title = element_text(size= 16, face = "bold"))
```

### Calculate the connectivity 
```{r Connectivity, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE}
rm(list = setdiff(ls(), c("Data", "GeneModules", "v_BatchMus")))

mainDir = "outputs/consensusWGCNA/"
subDir = "GeneConnectivity"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "/Connectivity_Consensus0.4.RData"))){
  Consensus <- "Consensus0.4"
  load(paste0("outputs/consensusWGCNA/dissTOM/dissTOM_", Consensus, ".RData"))
  consensusTOM <- 1 - dissTOM
  rm(dissTOM)
  filenames <- list.files(recursive = TRUE, path = "outputs/consensusWGCNA", pattern = paste0(Consensus, ".csv"), full.names = TRUE)
  # Importing gene-modules table
  GeneModules <- read.csv(filenames[grepl("All_genes_modules", filenames)]) %>%
    mutate(hgnc_symbol = v_BatchMus$genes$hgnc_symbol[match(NAMES, v_BatchMus$genes$ensembl_gene_id)])
  all(colnames(consensusTOM) == GeneModules$NAMES)
  COL <- GeneModules$mergedColors
  # Calculate the connectivity
  ConnectivityMat <- WGCNA::intramodularConnectivity(consensusTOM, COL)
  rm(consensusTOM)
  save(ConnectivityMat, file = paste0("outputs/consensusWGCNA/GeneConnectivity/Connectivity_", Consensus, ".RData"))
}
```

#### Hub genes in each module based on the within module connectivity
```{r HubGens, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5}
rm(list = setdiff(ls(), c("Data", "v_BatchMus")))

mainDir = "outputs/consensusWGCNA/"
subDir = "HubGenes"
if(!file.exists(mainDir, subDir)) dir.create(file.path(mainDir, subDir), recursive = TRUE)
if(!file.exists(paste0(mainDir, subDir, "/HubGenes_Consensus0.4.RData"))){
  # Top hub genes based on the kWithin
  load("outputs/consensusWGCNA/GeneConnectivity/Connectivity_Consensus0.4.RData")
  Consensus <- "Consensus0.4"

  filenames <- list.files(recursive = TRUE, path = "outputs/consensusWGCNA", pattern = paste0(Consensus, ".csv"), full.names = TRUE)

  # Importing gene-modules table
  GeneModules <- read.csv(filenames[grepl("All_genes_modules", filenames)]) %>%
    mutate(hgnc_symbol = v_BatchMus$genes$hgnc_symbol[match(NAMES, v_BatchMus$genes$ensembl_gene_id)])

  # Rename modules(colors to numbers)
  Module_rename <- data.frame(Module = prob::setdiff(sort(unique(GeneModules$mergedColors)), "grey"), 
                               Rename = paste("M", seq.int(length(unique(GeneModules$mergedColors))-1), sep = "."),
                               stringsAsFactors = FALSE) 

  GeneModules$mergedColors <- Module_rename$Rename[match(GeneModules$mergedColors, Module_rename$Module)] 
  GeneModules$mergedColors[is.na(GeneModules$mergedColors)] = "grey"

  modules <- unique(GeneModules$mergedColors)
  modules <- modules[modules != "grey"]
  hubs <- c()
	
  for(m in modules){
    GENES <- GeneModules %>% filter(mergedColors == m) %>% .$NAMES
		ConnectivitySub <- ConnectivityMat[rownames(ConnectivityMat) %in% GENES,] %>% 
		    arrange(- kWithin) 
		hubsM <- rownames(ConnectivitySub[1:20,])
		hubs[[m]] <- hubsM
 }
  save(hubs, file = paste0("outputs/consensusWGCNA/HubGenes/HubGenes_", Consensus, ".RData"))
}

```

```{r}
sessionInfo()
```