---
  title: "Capillary density"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

The muscle abbreviations were GRA (for Gracilis), SED (semitendinosus distal),  SEM (semitendinosus middle)), GAL (gastrocnemius lateralis), REF (rectus femoris), VAM (vastus medialis) and VAL (vastus lateralis\) in all the analyses and they only changed to GR, STM, STD, GL, RF, VM and VL for the publications.
```{r setup, include = FALSE}
rm (list = ls ())
suppressPackageStartupMessages ({ 
  library (data.table)
  library (dplyr)
  library (ggplot2)
  library (scales) 
  library (dplyr)
  library (plotly)
  library (knitr)
  library (cowplot)
  library (tidyr)
  library (grid)
  library (ggpubr)
  library (rgl)
  library(lmerTest)
})

knit_hooks$set(webgl = hook_webgl)
set.seed (1)

setwd("F://Corona-WorkingFromHome/Human-Muscle-Types/OurDataset/MuscleDAtaset/HumanMuscleTranscriptomeAtlasAnalyses/")
# Read all the txt files:
InputPath = "imageQuantification/input/capillaryDensity/Myofiber/"
Files <- list.files (path = InputPath, pattern = ".txt", full.names = T)
DataTable <- lapply (Files, read.table)
# Fixing a typo in image name file:
names (DataTable) <- gsub (":.*", "", sapply (DataTable, function (x) x[1, 1]))
DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s0`$Label <- gsub ("SED", "SEM", DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s0`$Label)
DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s1`$Label <- gsub ("SED", "SEM", DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s1`$Label)
DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s2`$Label <- gsub ("SED", "SEM", DataTable$`MD22_SED_Sam92-ECS_2021.03.24_s2`$Label)
names (DataTable)  <- gsub ("SED", "SEM", names (DataTable))
```

STM and GL sections of 20 individuals were stained with CD31 and ENG.
All the steps including segmentation were done automatically, but we do not need to do any filtration as we will consider the total area.

# Calculating the area
```{r Step2, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE, fig.height = 12, fig.width = 14}
FiberData <- rbindlist (DataTable, idcol = TRUE) %>% 
  select (.id, Label, Mean, Ch, Area) %>%
  unique () %>%
  mutate (Sample = gsub ("-.*", "", .id)) %>%
  mutate (Replicate = gsub (".*_", "", .id)) %>%
  mutate (Individual = gsub ("_.*", "", .id)) %>%
  mutate (Muscle = gsub ("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  mutate (fiber = gsub (":.$", "", Label))

FiberData <- tidyr::pivot_wider (FiberData [,c ("fiber", "Individual", "Muscle", "Replicate", "Ch", "Mean", "Area")], names_from = Ch, values_from = Mean) 

No_Area_Fibers <- merge (FiberData %>%
  count (Muscle, Individual, Replicate) %>%
  rename (NoFiber = n), FiberData %>%
    group_by (Muscle, Individual, Replicate) %>%
    summarise (AreaFiber = sum (Area)))

FinalSamples <- No_Area_Fibers %>% 
  group_by (Muscle, Individual) %>%
  top_n (1, NoFiber) %>% # Selecting the replicate with the highest number of myofibers
  filter (Individual != "MD19") %>% # The GL sample had a few number of myofibers so MD19 is excluded. 
  select (Muscle, Individual, Replicate)
```

# Improting quantification for capillaries
```{r EndoCells, include = FALSE}
rm (list = setdiff (ls (), c ("DataTable", "col.muscle", "FiberData", "NoFibers", "No_Area_Fibers", "FinalSamples")))

# Read all the txt files:
InputPath = "imageQuantification/input/capillaryDensity/ENG_CD31/"
Files <- list.files (path = InputPath, pattern = "ENG", full.names = T)

DataTable_ENG <- lapply (Files, read.table)
names (DataTable_ENG) <- gsub (".*/|_AnalyzeParticles.txt", "", Files)
names (DataTable_ENG)  <- gsub ("SED", "SEM", names (DataTable_ENG))

Files <- list.files (path = InputPath, pattern = "CD31", full.names = T)

DataTable_CD31 <- lapply (Files, read.table)
names (DataTable_CD31) <- gsub (".*/|_AnalyzeParticles.txt", "", Files)
names (DataTable_CD31)  <- gsub ("SED", "SEM", names (DataTable_CD31))

rm (list = setdiff (ls (), c ("DataTable", "col.muscle", "FiberData", "NoFibers", "No_Area_Fibers", "FinalSamples", "DataTable_CD31", "DataTable_ENG")))
```

# Filtering the capillaries and drawing the plots
```{r CD31ENG_Filtering, units = "cm", fig.width = 8, fig.height = 10}
DataToFilt_CD31 <- rbindlist (DataTable_CD31, idcol = TRUE) %>% 
  select (- Label) %>%
  unique () %>%
  mutate (Sample = gsub ("-.*", "", .id)) %>%
  mutate (Replicate = gsub ("_.*", "", gsub (".*_s", "s", .id))) %>%
  mutate (Individual = gsub ("_.*", "", .id)) %>%
  mutate (Muscle = gsub ("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  rename (id_CD31 = .id, Mean_CD31 = Mean, StdDev_CD31 = StdDev, Mode_CD31 = Mode,
          Min_CD31 = Min, Max_CD31 = Max, Median_CD31 = Median,
          XM_CD31 = XM, YM_CD31 = YM)

DataToFilt_ENG <- rbindlist (DataTable_ENG, idcol = TRUE) %>% 
  select (- Label) %>%
  unique () %>%
  mutate (Sample = gsub ("-.*", "", .id)) %>%
  mutate (Replicate = gsub ("_.*", "", gsub (".*_s", "s", .id))) %>%
  mutate (Individual = gsub ("_.*", "", .id)) %>%
  mutate (Muscle = gsub ("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  rename (id_ENG = .id, Mean_ENG = Mean, StdDev_ENG = StdDev, Mode_ENG = Mode,
          Min_ENG = Min, Max_ENG = Max, Median_ENG = Median,
          XM_ENG = XM, YM_ENG = YM)

# Organizing the tables
if (all (DataToFilt_ENG [ , c ("Area", "Circ.", "AR", "Round", "Solidity", "Sample", "Replicate", "Individual", "Muscle")] == DataToFilt_CD31 [, c ("Area", "Circ.", "AR", "Round", "Solidity", "Sample","Replicate", "Individual", "Muscle")])) {

DataToFilt <- cbind (DataToFilt_CD31, DataToFilt_ENG) %>% as.data.frame (as.is = T)
for (col in c ("Area", "Circ.", "AR", "Round", "Solidity", "Sample", "Replicate", "Individual", "Muscle")) {
  X = which (colnames (DataToFilt) == col)
  if (all (DataToFilt [, X[1]] == DataToFilt [, X[2]])) DataToFilt [, X[2]] = NULL
}
}

# Calculating the area positive for CD31
AreaCD31 <- DataToFilt %>%
  group_by (Muscle, Individual, Replicate) %>%
  summarise (AreaCD31 = sum (Area))

# %CD31 area
ORDER <- c ("GAL" = 2, "SEM" = 1)
col.muscle <- c ("GL" = "#999999", "STM" = "#0066FF") # Abbreviations based on the paper
ggplot (merge (merge (No_Area_Fibers, AreaCD31), FinalSamples) %>%
          mutate (Order = ORDER [match (Muscle, names (ORDER))]) %>%
          mutate (Muscle = ifelse (Muscle == "GAL", "GL", "STM")) %>%
          mutate (Pro_CD31Area = AreaCD31*100/AreaFiber),
        aes (x = reorder (Muscle, Order), y = Pro_CD31Area, fill = Muscle)) + 
  geom_boxplot (position = position_dodge(0.8), alpha = 0.5) +
  stat_summary (fun.y = mean, color = "darkred", position = position_dodge(0.75),
                geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual (values = col.muscle) +
  theme_bw () + labs (fill = element_blank(), y = "CD31 positive area (%)")  +
  theme (panel.grid = element_blank (), axis.text = element_text (size = 12), 
         axis.title.x = element_blank (), axis.title.y = element_text (size = 12), 
         axis.text.x.bottom = element_text (vjust = 0.5), axis.ticks = element_blank (),
         legend.position = "NULL")


flm <- as.formula ("Pro_CD31Area ~ 0 + Muscle + (1|Individual)")
MM <- lmerTest::lmer (flm,
                      # weights = Filt1_3,
                      data = merge (merge (No_Area_Fibers, AreaCD31), FinalSamples) %>%
                     mutate (Order = ORDER [match (Muscle, names (ORDER))]) %>%
                     mutate (Muscle = ifelse (Muscle == "GAL", "GL", "STM")) %>%
                     mutate (Pro_CD31Area = AreaCD31*100/AreaFiber))
print (anova (MM)["Muscle",]["Pr(>F)"])
#           Pr(>F)    
# Muscle 8.783e-09 ***
print (fixef (MM))
# MuscleGL MuscleSTM 
# 2.165429  1.358284

## Filtering 
DataToFilt_CircArea <- DataToFilt %>%
  filter (Circ. > 0.5) %>%
  filter (3 < Area) %>%
  filter (Area < 51)

DataToFilt$Included <- ifelse (paste (DataToFilt$id_CD31, DataToFilt$Area, DataToFilt$Circ., DataToFilt$Max_CD31, DataToFilt$Min_CD31) %in% paste (DataToFilt_CircArea$id_CD31, DataToFilt_CircArea$Area, DataToFilt_CircArea$Circ., DataToFilt_CircArea$Max_CD31, DataToFilt_CircArea$Min_CD31), 1, 0)

colnames (DataToFilt) <- gsub ("_CD31", "", colnames(DataToFilt))

# Keeping capillaries positive for both CD31 and ENG
DataToFilt_CircArea_ENG <- DataToFilt_CircArea %>%
  filter (Mean_ENG != 0)

# Number of capillaries
NoCapillaries <- merge (DataToFilt_CircArea %>%
                     count (Muscle, Individual, Replicate) %>%
                     rename (NoCD31 = n),
                     DataToFilt_CircArea_ENG %>%
                       count (Muscle, Individual, Replicate) %>%
                       rename (NoCD31_ENG = n))

# Table includes the number of capillaries before and after filtering
AreaFibers_NoCapillaries <- merge (merge (NoCapillaries, No_Area_Fibers) %>%
  mutate (ratio =  NoCD31*100/AreaFiber,
          Filtered_ratio = NoCD31_ENG*100/AreaFiber), FinalSamples) 

# Capilary density
ggplot (AreaFibers_NoCapillaries %>%
          mutate (Order = ORDER [match (Muscle, names (ORDER))]) %>%
          mutate (Muscle = ifelse (Muscle == "GAL", "GL", "STM")),
        aes (x = reorder (Muscle, Order), y = Filtered_ratio, fill = Muscle)) + 
  geom_boxplot (position = position_dodge(0.8), alpha = 0.5) +
  stat_summary (fun.y = mean, color = "darkred", position = position_dodge(0.75),
                geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual (values = col.muscle) +
  theme_bw () + labs (fill = element_blank(), y = "Capillary density")  +
  theme (panel.grid = element_blank (), axis.text = element_text (size = 12), 
         axis.title.x = element_blank (), axis.title.y = element_text (size = 12), 
         axis.text.x.bottom = element_text (vjust = 0.5), axis.ticks = element_blank (),
         legend.position = "NULL")

flm <- as.formula ("Filtered_ratio ~ 0 + Muscle + (1|Individual)")
MM <- lmerTest::lmer (flm,
                      # weights = Filt1_3,
                      data = AreaFibers_NoCapillaries)
print (anova (MM)["Muscle",]["Pr(>F)"])
#           Pr(>F)    
# Muscle 2.743e-09 ***
print (fixef (MM))
#  MuscleGAL  MuscleSEM 
# 0.03033244 0.02323773 

# logCPM CD31
Data <- new.env (); load ("dataAnalyses/output/batchMusCorrectedAllData.RData", envir = Data)
v_BatchMus <- Data$v_BatchMus
SubsetToPlot <- data.frame (CD31 = v_BatchMus$E [rownames (v_BatchMus) == "ENSG00000261371"],
                            # ENG = v_BatchMus$E [rownames (v_BatchMus) == "ENSG00000106991"],
                            Muscle = gsub ("MD[0-9][0-9]_|_Sam.*", "", colnames (v_BatchMus)),
                            Individual = gsub ("_.*", "", colnames (v_BatchMus)))

ToPlot <- merge (SubsetToPlot %>% 
                   reshape2::melt(value.name = "NormalizedExpression") %>%
                   rename (Gene = variable), 
                 AreaFibers_NoCapillaries, all.y = T)

ggplot (ToPlot %>% filter (Gene == "CD31") %>%
          mutate (Order = ORDER [match (Muscle, names (ORDER))]) %>%
          mutate (Muscle = ifelse (Muscle == "GAL", "GL", "STM"))%>%
          filter (Individual != "MD28"),
        aes (x = reorder (Muscle, Order), y = NormalizedExpression, fill = Muscle)) + 
  geom_boxplot (position = position_dodge(0.8), alpha = 0.5) +
  stat_summary (fun.y = mean, color = "darkred", position = position_dodge(0.75),
                geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual (values = col.muscle) +
  theme_bw () + labs (fill = element_blank(), y = expression("LogCPM"~italic ("CD31")))  +
  theme (panel.grid = element_blank (), axis.text = element_text (size = 12), 
         axis.title.x = element_blank (), axis.title.y = element_text (size = 12), 
         axis.text.x.bottom = element_text (vjust = 0.5), axis.ticks = element_blank (),
         legend.position = "NULL")
```


```{r}
sessionInfo ()
```