---
title: "RNAscope quantification"
author:
- tabbassidaloii
output:
 md_document:
    variant: markdown_github
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

'Note:' The muscle abbreviations were GRA(for Gracilis), SED(semitendinosus distal),  SEM(semitendinosus middle)), GAL(gastrocnemius lateralis), REF(rectus femoris), VAM(vastus medialis) and VAL(vastus lateralis) in all the analyses(scripts) and they were only changed to GR, STM, STD, GL, RF, VM and VL for the plots(shown in the publication).

```{r setup, include = FALSE}
rm(list = ls())

setwd(dirname(rstudioapi::getSourceEditorContext()$path))

if(!"dplyr" %in% installed.packages()) install.packages("dplyr")
if(!"data.table" %in% installed.packages()) install.packages("data.table")
if(!"ggplot2" %in% installed.packages()) install.packages("ggplot2")
if(!"tidyr" %in% installed.packages()) install.packages("tidyr")
if(!"cowplot" %in% installed.packages()) install.packages("cowplot")
if(!"lmerTest" %in% installed.packages())install.packages("lmerTest")
if(!"knitr" %in% installed.packages()) install.packages("knitr")
if(!"reshape2" %in% installed.packages()) install.packages("reshape2")
if(!"ggpubr" %in% installed.packages()) install.packages("ggpubr")

suppressPackageStartupMessages({
  library(dplyr)
  library(data.table)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(lmerTest)
  library(knitr)
  library(ggpubr)
  })

knit_hooks$set(webgl = hook_webgl)
set.seed(1)

# Read all the txt files:
InputPath = "input/RNAscope/AnalysisParticles/"
Files <- list.files(path = InputPath, pattern = ".txt", full.names = T, recursive = T)

# Remove image with no signal(MD19)
Files <- Files[-grep("MD19_GAL_H5_2Tiles_AnalyzeParticles", Files)]

DataTable_whole <- lapply(Files[grep("AnalyzeParticles.txt", Files)], read.csv, sep = "\t")
names(DataTable_whole) <- gsub(".*/|_AnalyzeParticles.txt| ", "", Files[grep("AnalyzeParticles.txt", Files)])

MyofiberSize <- lapply(Files[grep("ROIs.txt", Files)], read.csv, sep = "\t")
names(MyofiberSize) <- gsub(".*/|_ROIs.txt| ", "", Files[grep("ROIs.txt", Files)])
```

Muscles sections of 17 individuals were included in the RNAscope experiment. The samples were imaged and image processing and quantification were done in collaboration with Lennard Voortman. 

```{r ROI, echo = FALSE, message = FALSE, warning = FALSE, comment = F, fig.show = FALSE, fig.height = 12, fig.width = 14}
Data_ROI_size <- rbindlist(MyofiberSize, idcol = TRUE) %>%
  dplyr::select(-X) %>%
  rename(Sample = .id, FiberSize = Area, FiberX = XM, FiberY = YM, FiberCirc = Circ.) %>%
  mutate(ROI_pos = gsub(".*:", "", Label),
          Muscle = gsub("_.*", "", gsub("^[^_]*_", "", Sample)),
          Muscle = gsub("GAl", "GAL", Muscle), # correcting a typo in the files
          Individual = gsub("_.*", "", Sample),
          Type = gsub(".*_Neg.*", "Neg", gsub(".*_H.*", "Probes", Sample))) %>%
  dplyr::select(Sample, Individual, Muscle, Type, ROI_pos,
                 FiberSize, FiberCirc, FiberX, FiberY)

NoMyofiber <- Data_ROI_size %>%
  select(Sample:FiberY) %>%
  group_by(Individual, Muscle, Type) %>%
  count(name = "NoMyofibers")

Data_ROI <- rbindlist(DataTable_whole, idcol = TRUE) %>%
  dplyr::select(-X) %>%
  rename(Sample = .id, FociSize = Area, FociMFI = Mean, FociX = XM, FociY = YM, FociCirc = Circ.) %>%
  mutate(Channel = gsub("-.*", "", Label),
          ROI_pos = gsub(".*:", "", Label),
          Muscle = gsub("_.*", "", gsub("^[^_]*_", "", Sample)),
          Muscle = gsub("GAl", "GAL", Muscle),
          Individual = gsub("_.*", "", Sample),
          Type = gsub(".*_Neg.*", "Neg", gsub(".*_H.*", "Probes", Sample))) %>%
  dplyr::select(Sample, Individual, Muscle, Type, ROI_pos, 
                 Channel, FociSize, FociCirc, FociMFI, FociX, FociY)

Data_ROI$Channel[Data_ROI$Channel == "C2"] = "HOXA10"
Data_ROI$Channel[Data_ROI$Channel == "C4"] = "HOXC10"
Data_ROI$Channel[Data_ROI$Channel == "C5"] = "HOXA11"
    
Data_ROI <- merge(Data_ROI_size, Data_ROI, all = T, sort = F)

NoFoci <- merge(NoMyofiber, Data_ROI %>%
  group_by(Individual, Muscle, Type, Channel) %>%
  count(name = "NoFoci") %>%
  filter(!is.na(Channel)), all = T, by = c("Individual", "Muscle", "Type"))

## Whole ROI
#Foci Size
P1 <- ggplot(Data_ROI, aes_string(x = "FociSize")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("A: Foci Size") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(. ~ Type, scales='free_y', nrow = 2)

P2 <- ggplot(Data_ROI %>% filter(FociSize < 6), aes_string(x = "FociSize")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("B: Foci Size < 6um2") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(Type ~ Channel, scales='free_y', nrow = 2)

#Foci Circularity
P3 <- ggplot(Data_ROI, aes_string(x = "FociCirc")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("C: Foci Circ") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(. ~ Type, scales='free_y', nrow = 2)

P4 <- ggplot(Data_ROI %>% filter(FociCirc > 0.85), aes_string(x = "FociCirc")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("D: Foci Circ > .08") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(Type ~ Channel, scales='free_y', nrow = 2)

(Tosave <- plot_grid(P1, P2, P3, P4, hjust = 0, vjust = 1, ncol = 2, rel_widths = c(1.2, 3)))

Data_ROI_Filtered <- Data_ROI %>%
  filter(FociSize < 3.5 & FociCirc > 0.98) 
Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXC10"] <- ifelse(
  Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXC10"] < 75, 0, Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXC10"])
Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXA10"] <- ifelse(
  Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXA10"] < 300, 0, Data_ROI_Filtered$FociMFI[Data_ROI_Filtered$Channel == "HOXA10"])
Data_ROI_Filtered <- merge(Data_ROI %>% dplyr::select(Sample:FiberY) %>% unique(), 
                            Data_ROI_Filtered %>%
                              filter(FociMFI != 0), all = T)

(Data_ROI_Filtered %>% filter(!is.na(Channel)) %>% dim())[1] /(Data_ROI %>% filter(!is.na(Channel)) %>% dim())[1]

NoFoci_Filtered <- merge(NoFoci %>% select(-NoFoci), Data_ROI_Filtered %>%
  group_by(Individual, Muscle, Type, Channel) %>%
  count(name = "NoFoci") %>%
  filter(!is.na(Channel)), all = T)
NoFoci_Filtered$NoFoci[is.na(NoFoci_Filtered$NoFoci)] = 0

P1 <- ggplot(Data_ROI_Filtered, aes_string(x = "FociCirc")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("Foci Circ") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(. ~ Type, scales='free', nrow = 3)
P2 <- ggplot(Data_ROI_Filtered, aes_string(x = "FociSize")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("Foci Size") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(. ~ Type, scales='free', nrow = 3)
P3 <- ggplot(Data_ROI_Filtered, aes_string(x = "FociMFI")) +
    geom_density(alpha = .5,) +
    theme_bw() + ggtitle("FociMFI") +
    theme(legend.position = "none", axis.title = element_blank()) +
    facet_wrap(. ~ Type, scales='free_y', nrow = 2)
(Tosave <- plot_grid(P1, P2, P3, hjust = 0, vjust = 1, ncol = 3))

## Average number of foci per myofiber(unfiltered data)
Data_ROI_summary <- Data_ROI %>% 
  # filter(!is.na(Channel)) %>%
  group_by(Individual, Muscle, ROI_pos, Type, FiberSize, FiberCirc, Channel) %>%
  dplyr::summarise(NoFoci = n()) %>%
  ungroup()
Data_ROI_summary <- pivot_wider(Data_ROI_summary, names_from = Channel, values_from = NoFoci) 
Data_ROI_summary$`NA` <- NULL
Data_ROI_summary$HOXA10[is.na(Data_ROI_summary$HOXA10)] = 0
Data_ROI_summary$HOXA11[is.na(Data_ROI_summary$HOXA11)] = 0
Data_ROI_summary$HOXC10[is.na(Data_ROI_summary$HOXC10)] = 0
Data_ROI_summary <- pivot_longer(Data_ROI_summary, cols = c(HOXA10, HOXA11, HOXC10), 
                                  names_to = "Gene", values_to = "NoFoci")

# Association between myofiber size and number of foci
cor(Data_ROI_summary$FiberSize, Data_ROI_summary$NoFoci)
Data_ROI_Ave_summary <- merge(NoMyofiber, Data_ROI_summary %>% 
  group_by(Individual, Muscle, Type, Gene) %>%
  dplyr::summarise(Mean = mean(NoFoci),
                    Median = median(NoFoci)) %>%
  ungroup())

# For paper
col.muscle <- c("GL" = "#999999", "STM" = "#0066FF") # Abbreviations based on the paper

(P1 <- ggplot(Data_ROI_Ave_summary %>%
                mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = Gene, y = Mean, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8)) + 
  stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Average number of foci") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom") +
         facet_wrap(. ~ Type, nrow = 1))

## Average number of foci per myofiber(filtered data)
Data_ROI_Filtered_summary <- Data_ROI_Filtered %>% 
  group_by(Individual, Muscle, ROI_pos, Type, FiberSize, FiberCirc, Channel) %>%
  dplyr::summarise(NoFoci = n()) %>%
  ungroup()

Data_ROI_Filtered_summary <- pivot_wider(Data_ROI_Filtered_summary, names_from = Channel, values_from = NoFoci) 
Data_ROI_Filtered_summary$`NA` <- NULL
Data_ROI_Filtered_summary$HOXA10[is.na(Data_ROI_Filtered_summary$HOXA10)] = 0
Data_ROI_Filtered_summary$HOXA11[is.na(Data_ROI_Filtered_summary$HOXA11)] = 0
Data_ROI_Filtered_summary$HOXC10[is.na(Data_ROI_Filtered_summary$HOXC10)] = 0
Data_ROI_Filtered_summary <- pivot_longer(Data_ROI_Filtered_summary, 
                                           cols = c(HOXA10, HOXA11, HOXC10), 
                                  names_to = "Gene", values_to = "NoFoci")

(Tosave <- ggplot(Data_ROI_Filtered_summary %>%
          mutate(Sample = paste0(Individual, Muscle, Type),
                Muscle = ifelse(Muscle == "GAL", "GL", "STM")), 
        aes(x = NoFoci, group = Sample, color = Muscle)) +
   geom_density(adjust=1.5, alpha=.4) +
   theme_bw() + labs(x = "Number of foci per myofiber") +
   scale_color_manual(values = col.muscle) +
   facet_wrap(Type ~ Gene, scales = "free", nrow = 2))
  
 
(Tosave <- ggplot(Data_ROI_Filtered_summary,
        aes(x = Individual, y = NoFoci, fill = Gene)) + 
  geom_boxplot(position = position_dodge(0.8)) +
  # stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             # geom = "point", size = 3, show.legend = FALSE) +
  # scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Number of foci per myofiber") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom",
         axis.text.x.bottom = element_text(angle = 90, vjust = 0.5)) +
         facet_grid(Muscle ~ Type, scales = "free", space = "free"))

(Tosave <- ggplot(Data_ROI_Filtered_summary%>%
                mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = Individual, y = NoFoci, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8)) +
  # stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             # geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Number of foci per myofiber") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom",
         axis.text.x.bottom = element_text(angle = 90, vjust = 0.5)) +
         facet_grid(Gene ~ Type, scales = "free", space = "free"))

(Tosave <- ggplot(Data_ROI_Filtered_summary %>%
                     filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
                     filter(Gene != "HOXA11")%>%
                mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = Individual, y = NoFoci, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8)) +
  # stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             # geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Number of foci per myofiber") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom",
         axis.text.x.bottom = element_text(angle = 90, vjust = 0.5)) +
         facet_grid(Gene ~ Type, scales = "free", space = "free"))

# Association between myofiber size and number of foci
cor(Data_ROI_Filtered_summary$FiberSize, Data_ROI_Filtered_summary$NoFoci)
Data_ROI_Filtered_Ave_summary <- merge(NoMyofiber, Data_ROI_Filtered_summary %>% 
  group_by(Individual, Muscle, Type, Gene) %>%
  dplyr::summarise(Mean = mean(NoFoci),
                    Median = median(NoFoci)) %>%
  ungroup(), all = T)

P2 <- ggplot(Data_ROI_Filtered_Ave_summary%>%
                mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = Gene, y = Mean, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8)) +
  stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Average number of foci") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom") +
         facet_wrap(. ~ Type, nrow = 1)
(Tosave <- plot_grid(P1, P2, hjust = 0, vjust = 1, ncol = 1, labels = c("Before filtering", "After filtering")))

(Tosave <- ggplot(Data_ROI_Filtered_Ave_summary %>%
                     filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
                     filter(Gene != "HOXA11")%>%
                mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = Gene, y = Mean, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8)) +
  stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Average number of foci") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 12), 
         axis.title.x = element_blank(), legend.position = "bottom") +
         facet_wrap(. ~ Type, nrow = 1))

ORDER <- c("GL" = 2, "STM" = 1)

(Tosave <- ggplot(Data_ROI_Filtered_Ave_summary %>%
                     filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
                     filter(Gene != "HOXA11") %>%
                     filter(Type != "Neg") %>%
                     mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM"))%>%
                     mutate(Order = ORDER[match(Muscle, names(ORDER))]),
        aes(x = reorder(Muscle, Order), y = Mean, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8), alpha = 0.5) +
   stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Average number of foci") +
  theme(panel.grid = element_blank(), axis.text = element_text(size = 12), 
         axis.title.x = element_blank(), axis.title.y = element_text(size = 12),
         axis.text.x.bottom = element_text(vjust = 0.5), axis.ticks = element_blank(),
         legend.position = "NULL", strip.text = element_text(size = 12, face = "bold.italic")) +
    facet_wrap(~ Gene, nrow = 2))
ggsave(Tosave, device = "tiff", units = "cm", width = 8, height = 11, filename = "figures/HOX_Foci.tif")

ToStat <- Data_ROI_Filtered_Ave_summary %>%
                     filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
                     filter(Gene != "HOXA11") %>%
                     filter(Type != "Neg") %>%
                     mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM"))%>%
                     mutate(Order = ORDER[match(Muscle, names(ORDER))])

flm <- as.formula("Mean ~ 0 + Muscle +(1|Individual)")
MM <- lmerTest::lmer(flm,
                      data = ToStat %>% filter(Gene == "HOXA10"))
print(anova(MM)["Muscle",]["Pr(>F)"])
print(fixef(MM))
MM <- lmerTest::lmer(flm,
                      data = ToStat %>% filter(Gene == "HOXC10"))
print(anova(MM)["Muscle",]["Pr(>F)"])
print(fixef(MM))

Data <- new.env(); load("../dataAnalyses/outputs/correctedGenesDatasetToAnalysis.RData", envir = Data)
v_BatchMus <- Data$v_BatchMus
SubsetToPlot <- data.frame(HOXA10 = v_BatchMus$E[rownames(v_BatchMus) == "ENSG00000253293"],
                            HOXA11 = v_BatchMus$E[rownames(v_BatchMus) == "ENSG00000005073"],
                            HOXC10 = v_BatchMus$E[rownames(v_BatchMus) == "ENSG00000180818"],
                            Muscle = gsub("MD[0-9][0-9]_|_Sam.*", "", colnames(v_BatchMus)),
                            Individual = gsub("_.*", "", colnames(v_BatchMus)))
ToPlot <- merge(SubsetToPlot %>% 
                   reshape2::melt(value.name = "NormalizedExpression") %>%
                   rename(Gene = variable), 
                 Data_ROI_Ave_summary %>% 
                   filter(Type == "Probes")) %>%
  filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
  filter(Gene != "HOXA11")

(Tosave <- ggplot(ToPlot, aes(x = NormalizedExpression, y = Mean, color = Muscle)) +
  geom_point(size = 1.5) +
  scale_color_manual(values = col.muscle) +
  geom_smooth(method = lm, colour = "black") + stat_cor(method = "pearson") + theme_bw() +
  labs(y = "Average number of foci", x = "Normalized expression", color = "") +
  theme(panel.grid = element_blank(),axis.title =  element_text(size = 14),
         axis.text.x = element_text(size = 12, vjust = 0.5),
         # axis.ticks = element_blank(),
         legend.text = element_text( size = 12, vjust = 0.5),
         strip.text = element_text(size = 12, face = "bold")) +
  facet_wrap(. ~ Gene, nrow = 3, scales = "free"))

ggsave(Tosave, device = "tiff", units = "cm", width = 9, height = 12, filename = "figures/HOX_Foci_Gen.tif")


(Tosave <- ggplot(ToPlot %>%
                     filter(!Individual %in% c("MD28")) %>%
                     filter(Gene != "HOXA11") %>%
                     filter(Type != "Neg") %>%
                     mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")) %>%
                     mutate(Order = ORDER[match(Muscle, names(ORDER))]),
        aes(x = reorder(Muscle, Order), y = NormalizedExpression, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8), alpha = 0.5) +
    stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
             geom = "point", size = 3, show.legend = FALSE) +
    scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "LogCPM") +
  theme(panel.grid = element_blank(), axis.text = element_text(size = 12), 
         axis.title.x = element_blank(), axis.title.y = element_text(size = 12),
         axis.text.x.bottom = element_text(vjust = 0.5), axis.ticks = element_blank(),
         legend.position = "NULL", strip.text = element_text(size = 12, face = "bold.italic")) +
    facet_wrap(~ Gene, nrow = 2))
ggsave(Tosave, device = "tiff", units = "cm", width = 8, height = 11, filename = "figures/HOX_Gen.tif")

Data_ROI_Filtered_Ave_summary <- Data_ROI_Filtered_Ave_summary %>%
  select(-Median) %>%
  pivot_wider(names_from = Gene, values_from = Mean) 

comparison <- c()
for(Gene in c("HOXA10", "HOXA11", "HOXC10")){
    flm <- as.formula(paste(noquote(Gene), "~ 0 + Muscle +(1|Individual)"))
    MM <- lmerTest::lmer(flm, weights = NoMyofibers,
                              data = Data_ROI_Filtered_Ave_summary %>% filter(Type == "Probes"))
    print(Gene)
    print(anova(MM)["Muscle",]["Pr(>F)"])
    print(fixef(MM))
    x <- data.frame(Gene, anova(MM)["Muscle",]["Pr(>F)"], fixef(MM)["MuscleSEM"], fixef(MM)["MuscleGAL"])
    comparison <- rbind(comparison, x)
}
colnames(comparison)[2:4] <- c("Pr(>F)", "Mean_SEM", "Mean_GAL") 
rownames(comparison) <- NULL

Data_ROI_Filtered_Ave_summary <- Data_ROI_Filtered_Ave_summary %>%
  filter(!Individual %in% c("MD13", "MD14", "MD17", "MD22", "MD27")) %>%
  filter(Gene != "HOXA11")
comparison <- c()
for(Gene in c("HOXA10", "HOXC10")){
    flm <- as.formula(paste(noquote(Gene), "~ 0 + Muscle +(1|Individual)"))
    MM <- lmerTest::lmer(flm, weights = NoMyofibers,
                              data = Data_ROI_Filtered_Ave_summary %>% filter(Type == "Probes"))
    print(Gene)
    print(anova(MM)["Muscle",]["Pr(>F)"])
    print(fixef(MM))
    x <- data.frame(Gene, anova(MM)["Muscle",]["Pr(>F)"], fixef(MM)["MuscleSEM"], fixef(MM)["MuscleGAL"])
    comparison <- rbind(comparison, x)
}
colnames(comparison)[2:4] <- c("Pr(>F)", "Mean_SEM", "Mean_GAL")
rownames(comparison) <- NULL
#     Gene       Pr(>F) Mean_SEM Mean_GAL
# 1 HOXA10 2.582160e-07 5.661107  4.54487
# 2 HOXC10 1.310777e-07 4.565500  3.42333

```